### 2022-06-23 data/Gu-thesis-script.Rmd
First draft

---
title: 'parental migration and adolescents educational performance'
output: html_notebook
---

Load packages
```{r}
library(tidyverse) 
library(haven)
library(fixest)
library(modelsummary)
library(MatchIt)
#library(vtable)
library(ggplot2)
#library(gtsummary)
#library(DescTools)
#knitr::opts_chunk$set(modelsummary_get="broom")
```

## Data cleaning
Read CEPS data
```{r}
d_stu1 <- read_dta('cepsw1studentCN.dta') #wave 1 student
d_stu2 <- read_dta('cepsw2studentCN.dta') #wave 2 student
d_par1 <- read_dta('cepsw1parentCN.dta') #wave 1 parent
d_par2 <- read_dta('cepsw2parentCN.dta') #wave 2 parent
#d_sch1 <- read_dta('cepsw1principalCN.dta') #wave 1 school principal
#d_sch2 <- read_dta('cepsw2principalCN.dta')
```

Remove wave 1's observations of grade 9; they were not followed up in wave 2
```{r}
d_stu1 <- d_stu1 %>% 
  subset(grade9==0) %>% 
  select(!c(frame, subsample, sweight, grade9))
d_stu2 <- d_stu2 %>% 
  select(!c(w2frame, w2subsample, w2sweight, w1w2sweight))
d_par1 <- d_par1 %>% 
  subset(grade9==0) %>% 
  select(!c(frame, subsample, grade9))
```

Assemble a panel dataset in wide format
```{r}
dw1 <- d_stu1 %>% left_join(d_par1, by=c('ids', 'schids', 'clsids', 'ctyids', 'fall')) 
dw2 <- d_stu2 %>% left_join(d_par2, by=c('ids', 'clsids', 'w2clsids', 'schids', 'ctyids', 'w2fall'))
d <- dw1 %>% right_join(dw2, by=c('ids', 'clsids', 'schids', 'ctyids'))
```

```{r}
d <- d %>% mutate(
  child.age = 2013 - a02a, #student's age at baseline
  w1chn_100 = tr_chn/w2upchn*100, #raw exam scores at 100-point scale
  w1mat_100 = tr_mat/w2upmat*100,
  w1eng_100 = tr_eng/w2upeng*100,
  w2chn_100 = w2chn/w2upchn*100,
  w2mat_100 = w2mat/w2upmat*100,
  w2eng_100 = w2eng/w2upeng*100
)
```

```{}
set.seed(1)
d %>% select(
  w1chn_100:w2eng_100) %>%
ggplot(aes(x=w1chn_100:w2eng_100)) + 
  geom_boxplot(alpha=0.2)
```

Trimming outliers
```{r}
d <- d %>% mutate(
  w1chn_100_trim = if_else(w1chn_100>100, 100, w1chn_100),
  w2chn_100_trim = if_else(w2chn_100>100, 100, w2chn_100),
  w1mat_100_trim = if_else(w1mat_100>100, 100, w1mat_100),
  w2mat_100_trim = if_else(w2mat_100>100, 100, w2mat_100),
  w1eng_100_trim = if_else(w1eng_100>100, 100, w1eng_100),
  w2eng_100_trim = if_else(w2eng_100>100, 100, w2eng_100)
)
```

```{r}
d <- d %>% mutate(
  w1ttl_100 = w1chn_100_trim + w1mat_100_trim + w1eng_100_trim,
  w2ttl_100 = w2chn_100_trim + w2mat_100_trim + w2eng_100_trim,
  ttl_dif = w2ttl_100 - w1ttl_100,
  cog_dif = w2cog3pl - cog3pl
  )
```

```{r}
d1 <- d %>%
  drop_na(cog_dif, ttl_dif, child.age)
```

within-class z score
```{}
d1 <- d1 %>% group_by(clsids + w2clsids) %>%
  mutate(
    w1ttl_z = as.numeric(scale(w1ttl)),
    w2ttl_z = as.numeric(scale(w2ttl))
  ) %>% ungroup()
```

```{}
d1 <- d1 %>% group_by(clsids + w2clsids) %>%
  mutate(
    w1ttl_std = rescale(w1ttl),
    w2ttl_std = rescale(w2ttl)
    ) %>% ungroup()
```

Detect outliers
```{}
d_outlier <- d %>% filter(w1chn_100 > 100 | w1mat_100>100|w1eng_100>100|w2chn_100>100|w2mat_100>100|w2eng_100>100)
d_outlier1 <- d_outlier %>% select(clsids, clsids + w2clsids, tr_chn, tr_mat, tr_eng, w2chn, w2upchn, w2mat, w2upmat, w2eng, w2upeng)
d1 <- d %>% anti_join(d_outlier, by=c('ids', 'clsids', 'clsids + w2clsids', 'ctyids'))
```

Sample
```{r}
d_cl <- d1 %>% 
  filter(w2a0801==1) %>% #keep parents who are married (and alive)
  filter(be1201==1 & be1202==1) %>% #keep both parents live with children at baseline
  filter(stmigrant==1) #keep local children 
```

## Variables

Control variables
```{r}
d_cl <- d_cl %>% mutate(
         child.is.girl = if_else(a01 == 2, 1, 0), #student is girl
         child.has.rural.hukou = sthktype, #student holds rural (agricultural) hukou
         child.in.boarding.school = stbrd, #student is boarding at school
         child.is.ethnic.minority = if_else(a03 != 1, 1, 0), #student is ethnic minority
         child.health.self.rated = a17, #student's self-rated health
         child.had.skipped.grade = if_else(c03 == 0, 0, 1), #student had skipped grade
         child.had.repeated.grade = if_else(c04 == 0, 0, 1), #student had repeated grade
         #ch.illness = if_else(a10 == 1, 1, 0), #student had serious illness before primary school
         child.went.to.preschool = if_else(c01 == 1, 1, 0), #student had attended preschool
         #ch.par.rltn = if_else(stprrel == 2, 1, 0),
         #ch.tutoring = if_else(b1901 == 1|b1902 == 1|b1903 == 1|b1904 == 1, 1, 0),
         par.edu = stprhedu, #Among two parents, the best-educated one's educational level
         parent.has.white.collar.job = if_else(be08==1|be08==2|be08==3, 1, 0),
         #par.ccp = if_else(be06 == 1, 1, 0), #At least one parent is communist party member
         home.economic.resource = steco_5c, #self-rated household financial situation
         home.has.internet = if_else(b13 == 2, 1, 0), #student has internet access at home
         home.extracurricular.book = b12 #Amount of book at home, student self-rated 
) %>% drop_na(
child.is.girl, child.has.rural.hukou, child.health.self.rated, child.in.boarding.school, child.is.ethnic.minority, child.health.self.rated, child.had.skipped.grade, child.had.repeated.grade, child.went.to.preschool, par.edu, 
#par.ccp, 
parent.has.white.collar.job, home.has.internet, home.extracurricular.book, home.economic.resource)
```

```{r}
d_cl$stsib <- d_cl$stsib %>% replace_na(0) %>% as.numeric()
d_cl <- d_cl %>% mutate(child.number.of.sibling = if_else(stonly == 1, 0, stsib)) %>% #number of siblings
  drop_na(child.number.of.sibling)
```

Best-educated parent's years of education
```{}
d_cl <- d_cl %>% mutate(parent.educational.level = NA)
d_cl$parent.educational.level[d_cl$par.edu == 1] <- 0
d_cl$parent.educational.level[d_cl$par.edu == 2] <- 6
d_cl$parent.educational.level[d_cl$par.edu == 3] <- 9
d_cl$parent.educational.level[d_cl$par.edu == 4|d_cl$par.edu == 5|d_cl$par.edu == 6] <- 12
d_cl$parent.educational.level[d_cl$par.edu == 7] <- 15
d_cl$parent.educational.level[d_cl$par.edu == 8] <- 16
d_cl$parent.educational.level[d_cl$par.edu == 9] <- 19
```

```{r}
d_cl <- d_cl %>% mutate(parent.educational.level = NA)
d_cl$parent.educational.level[d_cl$par.edu == 1|d_cl$par.edu == 2] <- 1
d_cl$parent.educational.level[d_cl$par.edu == 3] <- 2
d_cl$parent.educational.level[d_cl$par.edu == 4|d_cl$par.edu == 5|d_cl$par.edu == 6] <- 3
d_cl$parent.educational.level[d_cl$par.edu == 7] <- 4
d_cl$parent.educational.level[d_cl$par.edu == 8] <- 5
d_cl$parent.educational.level[d_cl$par.edu == 9] <- 6
```

```{r}
d_cl <- d_cl %>% mutate(parent.expectation.on.child = NA)
d_cl$parent.expectation.on.child[d_cl$ba18 == 1|d_cl$ba18 == 2] <- 1
d_cl$parent.expectation.on.child[d_cl$ba18 == 3|d_cl$ba18 == 4|d_cl$ba18 == 5] <- 2
d_cl$parent.expectation.on.child[d_cl$ba18 == 6] <- 3
d_cl$parent.expectation.on.child[d_cl$ba18 == 7] <- 4
d_cl$parent.expectation.on.child[d_cl$ba18 == 8|d_cl$ba18 == 9] <- 5
```

```{r}
d_cl <- d_cl %>% mutate(child.educational.aspiration = NA)
d_cl$child.educational.aspiration[d_cl$c22 == 1|d_cl$c22 == 2|d_cl$c22 == 10] <- 1
d_cl$child.educational.aspiration[d_cl$c22 == 3|d_cl$c22 == 4|d_cl$c22 == 5] <- 2
d_cl$child.educational.aspiration[d_cl$c22 == 6] <- 3
d_cl$child.educational.aspiration[d_cl$c22 == 7] <- 4
d_cl$child.educational.aspiration[d_cl$c22 == 8|d_cl$c22 == 9] <- 5
```

```{r}
d_cl <- d_cl %>%  drop_na(
  parent.educational.level, 
  parent.expectation.on.child, 
  child.educational.aspiration)
```

## Treatment groups
At least one parent migrated (Control group as no parent migrated)
```{r}
d_cl <- d_cl %>% mutate(
  migr = if_else((w2ba0201==0 | w2ba0202==0), 1, 0), #Any parent became absent from household between baseline and endline survey
  migr_fa_only = NA,
  #migr_fa = NA,
  migr_mo_only = NA,
  #migr_mo = NA,
  migr_both = NA
  ) %>% drop_na(migr)
```

Only mother migrated
```{r}
d_cl$migr_mo_only[d_cl$migr == 0] <- 0
d_cl$migr_mo_only[d_cl$w2ba0201==1 & d_cl$w2ba0202==0] <- 1 
```

Mother migrated
```{}
d_cl$migr_mo[d_cl$migr == 0] <- 0
d_cl$migr_mo[d_cl$w2ba0202==0] <- 1
```

Only father migrated
```{r}
d_cl$migr_fa_only[d_cl$migr == 0] <- 0
d_cl$migr_fa_only[d_cl$w2ba0201==0 & d_cl$w2ba0202==1] <- 1 
```

Father migrated
```{}
d_cl$migr_fa[d_cl$migr == 0] <- 0
d_cl$migr_fa[d_cl$w2ba0201==0] <- 1
```

Both parents migrated
```{r}
d_cl$migr_both[d_cl$migr == 0] <- 0
d_cl$migr_both[d_cl$w2ba0201==0 & d_cl$w2ba0202==0] <- 1
```

```{}
df <- d_cl %>% select(
  migr, migr_fa_only, migr_fa, migr_mo_only, migr_mo, migr_both,
  cog3pl, w1ttl_100, cog_dif, ttl_dif,
  child.is.girl, child.has.rural.hukou, child.health.self.rated, child.in.boarding.school, child.is.ethnic.minority, child.health.self.rated, child.had.skipped.grade, child.had.repeated.grade, child.went.to.preschool, child.educational.aspiration, 
  par.edu, par.ccp, par.expt.univ, 
  home.has.internet, home.extracurricular.book, home.economic.resource
)
```

## Descriptive statistics
```{r}
d_migr <- d_cl %>%
  select(
  migr, migr_mo_only, migr_fa_only, migr_both,
  cog3pl, w2cog3pl, w1ttl_100, w2ttl_100, cog_dif, ttl_dif,
  child.age, child.is.girl, child.has.rural.hukou, child.in.boarding.school, child.is.ethnic.minority, child.had.skipped.grade, child.had.repeated.grade, child.went.to.preschool, parent.has.white.collar.job, home.has.internet,
  child.health.self.rated, child.educational.aspiration, 
  parent.educational.level, parent.expectation.on.child, 
  home.extracurricular.book, home.economic.resource) %>%
  remove_labels()
```

```{r}
d_smr <- d_migr %>% 
  mutate(`Parental Migration` = if_else((migr==1), 'Yes', 'No')) %>% 
  mutate(`Parental Migration Type` = 0) 
d_smr$`Parental Migration Type`[d_smr$migr==0] <- 'Both Parents At Home'
d_smr$`Parental Migration Type`[d_smr$migr_mo_only==1] <- 'Only Mother Migrated'
d_smr$`Parental Migration Type`[d_smr$migr_fa_only==1] <- 'Only Father Migrated'
d_smr$`Parental Migration Type`[d_smr$migr_both==1] <- 'Both Parents Migrated'
d_smr <- d_smr %>% select(!migr & !migr_mo_only & !migr_fa_only & !migr_both)
```

```{r}
datasummary_balance(
  ~migr,
  data = d_migr,
  fmt = 2)
```

```{r}
d_dpvar <- d_migr %>%
  select(#migr, 
         cog_dif, cog3pl, w2cog3pl, ttl_dif, w1ttl_100, w2ttl_100) %>%
  rename(
    `Cognitive score at wave 1` = cog3pl,
    `Cognitive score at wave 2` = w2cog3pl,
    `Academic exam score at wave 1` = w1ttl_100,
    `Academic exam score at wave 2` = w2ttl_100,
    `Cognitive score difference` = cog_dif,
    `Academic exam score difference` = ttl_dif
  ) 
```

```{}
d_dpvar %>%
  tbl_summary(
    by = migr,
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} / {N} ({p}%)"),
    digits = all_continuous() ~ 2,
    ) %>% add_p() %>% as_kable()
```

```{r}
#table_independent_variable <- 
  st(
  d_dpvar,
  digits = 2, 
  group = 'migr',
  group.test = list(star.cutoffs = c(0.001, 0.01, 0.05)),
  out = 'kable')
#table_independent_variable 
```

```{r}
d_dpvar <- as.data.frame(d_dpvar)
table_dependent_variable <- datasummary(
  All(d_dpvar) ~ Mean + SD + Min + Max,
  data = d_dpvar,
  output = 'markdown'
  )
table_dependent_variable
```

```{r}
table_independent_variable <- datasummary_skim(
    data=d_idpvar,
    type = 'categorical',
    output = 'markdown')
table_independent_variable
```

```{r}
d_idpvar <- d_migr %>% 
  select(migr, migr_mo_only, migr_fa_only, migr_both) %>% 
  mutate(`Parental Migration` = if_else((migr==1), 'Yes', 'No')) %>% 
  mutate(`Parental Migration Type` = 0) 
d_idpvar$`Parental Migration Type`[d_idpvar$migr==0] <- 'Both Parents At Home'
d_idpvar$`Parental Migration Type`[d_idpvar$migr_mo_only==1] <- 'Only Mother Migrated'
d_idpvar$`Parental Migration Type`[d_idpvar$migr_fa_only==1] <- 'Only Father Migrated'
d_idpvar$`Parental Migration Type`[d_idpvar$migr_both==1] <- 'Both Parents Migrated'
d_idpvar <- d_idpvar %>% select(!migr & !migr_mo_only & !migr_fa_only & !migr_both)
```

```{r}

```

```{r}
d_ctvar <- d_migr %>%
  select(migr, child.is.girl, child.has.rural.hukou, child.in.boarding.school, child.is.ethnic.minority, child.had.skipped.grade, child.had.repeated.grade, child.went.to.preschool, parent.has.white.collar.job, home.has.internet,
  child.health.self.rated, child.educational.aspiration, 
  parent.educational.level, parent.expectation.on.child, 
  home.extracurricular.book, home.economic.resource) %>% 
  mutate_if(is.numeric, as.factor)
```

```{r}
d_ctvar$parent.educational.level <- d_ctvar$parent.educational.level %>% 
  recode_factor(
    '1' = 'Primary school and below',
    '2' = 'Junior high school',
    '3' = 'Senior high school',
    '4' = 'Associate college',
    '5' = 'University',
    '6' = 'Postgraduate'
  )
d_ctvar$child.educational.aspiration <- d_ctvar$child.educational.aspiration %>%
  recode_factor(
    '1' = 'Junior high school',
    '2' = 'Senior high school',
    '3' = 'Associate college',
    '4' = 'University',
    '5' = 'Postgraduate'     
  )
d_ctvar$parent.expectation.on.child  <- d_ctvar$parent.expectation.on.child %>%
  recode_factor(
    '1' = 'Junior high school',
    '2' = 'Senior high school',
    '3' = 'Associate college',
    '4' = 'University',
    '5' = 'Postgraduate'    
  )
d_ctvar$home.economic.resource <- d_ctvar$home.economic.resource %>% 
  recode_factor(
    '1' = 'Very poor',
    '2' = 'Somewhat poor',
    '3' = 'Moderate',
    '4' = 'Somewhat rich',
    '5' = 'Very rich'
  )
d_ctvar$child.health.self.rated <- d_ctvar$child.health.self.rated %>%
  recode_factor(
    '1' = 'Very poor',
    '2' = 'Not very good',
    '3' = 'Moderate',
    '4' = 'Good',
    '5' = 'Very good'
  )
d_ctvar$home.extracurricular.book <- d_ctvar$home.extracurricular.book %>%
  recode_factor(
    '1' = 'Very few',
    '2' = 'Not many',
    '3' = 'Some',
    '4' = 'Quite a few',
    '5' = 'A great number'
  )
```

```{}
test.table <- d_ctvar %>%
  tbl_summary(
    by = migr,
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
    digits = all_continuous() ~ 2,
    ) 
```

```{r}
test_table <- datasummary_balance(
  ~migr,
  data = d_ctvar
)
test_table
```

```{r}
#table_control_variable <- 
table_balance <- st(
  d_ctvar,
  digits = 2, 
  group = 'migr',
  group.test = list(star.cutoffs = c(0.001, 0.01, 0.05)),
  out = "return")
#table_control_variable 
```

```{r}
table_control_variable <- datasummary_skim(
  data = d_ctvar,
  type = 'categorical',
  output = 'markdown'
)
```

```{r}
d_migr <- d_migr %>% mutate(`Parental Migration` = ifelse(migr == 1, 'Yes', 'No'))
d_migr <- d_migr %>% mutate(`Parental Migration Type` = 0) 
d_migr$`Parental Migration Type`[d_migr$migr==0] <- 'Both Parents At Home'
d_migr$`Parental Migration Type`[d_migr$migr_mo_only==1] <- 'Only Mother Migrated'
d_migr$`Parental Migration Type`[d_migr$migr_fa_only==1] <- 'Only Father Migrated'
d_migr$`Parental Migration Type`[d_migr$migr_both==1] <- 'Both Parents Migrated'
d_migr <- d_migr %>% select(!migr & !migr_mo_only & !migr_fa_only & !migr_both)
```

Whole sample summary
```{r}
table_sum <- st(
  d_migr,
  digits = 2, 
  out = 'kable')
table_sum
```

```{r}
table_sum_migr <- st(
  d_migr,
  group = 'Parental Migration', 
  digits = 2, 
  group.test = list(star.cutoffs = c(0.001, 0.01, 0.05)),
  out = 'kable')
table_sum_migr
```

t-test
```{r}
ttest_cog <- t.test(cog_dif ~ migr, data = d_cl)
ttest_ttl <- t.test(ttl_dif ~ migr, data = d_cl)
ttest_cog
ttest_ttl
```

```{r}
set.seed(1)
fig_cog <- 
  ggplot(d_migr, aes(x=`Parental Migration`, y=cog_dif, fill = `Parental Migration`)) + 
    geom_violin()+
    xlab('Parental Migration') +
    theme(legend.position="none") +
    ylab("")
fig_cog
```

```{r}
set.seed(1)
fig_ttl <- 
  ggplot(d_migr, aes(x=`Parental Migration`, y=ttl_dif, fill = `Parental Migration`)) + 
    geom_violin()+
    xlab('Parental Migration') +
    theme(legend.position="none") +
    ylab("")
fig_ttl
```

```{r}
set.seed(1)
fig_cog_by_migration_type <-
  ggplot(d_migr, aes(x=`Parental Migration Type`, y=cog_dif, fill = `Parental Migration Type`)) + 
    geom_violin()+
    xlab('Parental Migration Type') +
    theme(legend.position="none") +
    ylab("")
fig_cog_by_migration_type
```

```{r}
set.seed(1)
fig_ttl_by_migration_type <-
  ggplot(d_migr, aes(x=`Parental Migration Type`, y=ttl_dif, fill = `Parental Migration Type`)) + 
    geom_violin()+
    xlab('Parental Migration Type') +
    theme(legend.position="none") +
    ylab("")
fig_ttl_by_migration_type
```

## DID unrestricted
Regressing on cognitive test score (standardized)
```{r}
did_cog <- feols(cog_dif ~ migr + cog3pl | clsids + w2clsids, data = d_cl) 
did_cog_m_only <- feols(cog_dif ~ migr_mo_only + cog3pl | clsids + w2clsids, data = d_cl) 
#did_cog_m <- feols(cog_dif ~ migr_mo + cog3pl | clsids + w2clsids, data = d_cl) 
did_cog_f_only <- feols(cog_dif ~ migr_fa_only + cog3pl | clsids + w2clsids, data = d_cl) 
#did_cog_f <- feols(cog_dif ~ migr_fa + cog3pl | clsids + w2clsids, data = d_cl) 
did_cog_both <- feols(cog_dif ~ migr_both + cog3pl | clsids + w2clsids, data = d_cl)
```

Results
```{r}
models_cog <- list(
'Any Parent Absent' = did_cog,
'Only Mother Absent' = did_cog_m_only,
#'Mother Absent' = #did_cog_m,
'Only Father Absent' = did_cog_f_only,
#'Father Absent' = #did_cog_f,
'Both Parents Absent' = did_cog_both) 
```

```{r}
stars <- c('†' = .1, '*' = .05, '**' = .01, '***' = .001)
coef_rename <- c("migr" = "Explanator", "migr_mo_only" = "Explanator", "migr_mo" = "Explanator", "migr_fa" = "Explanator", "migr_fa_only" = "Explanator", "migr_both" = "Explanator", "cog3pl" = "W1 score", "w1ttl_100" = "W1 score")
title_cog <- 'Parental migration’s effect on children’s cognitive abilities'
title_ttl <- 'Parental migration’s effect on children’s academic abilities'
gof_omit <- 'R2 Pseudo|AIC|BIC|Log.Lik.|Std.Errors'
```

```{r}
table_cog <- models_cog %>% msummary(
  stars =  stars,
  fmt = 2,
  title = title_cog,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  output = 'markdown')
table_cog
```

raw score
```{r}
did_ttl <- feols(ttl_dif ~ migr + w1ttl_100 | clsids + w2clsids, data = d_cl) 
did_ttl_m_only <- feols(ttl_dif ~ migr_mo_only + w1ttl_100 | clsids + w2clsids, data = d_cl) 
#did_ttl_m <- feols(ttl_dif ~ migr_mo + w1ttl_100 | clsids + w2clsids, data = d_cl) 
did_ttl_f_only <- feols(ttl_dif ~ migr_fa_only + w1ttl_100 | clsids + w2clsids, data = d_cl) 
#did_ttl_f <- feols(ttl_dif ~ migr_fa + w1ttl_100 | clsids + w2clsids, data = d_cl) 
did_ttl_both <- feols(ttl_dif ~ migr_both + w1ttl_100 | clsids + w2clsids, data = d_cl)
```

```{r}
models_ttl <- list(
'Any Parent Absent' = did_ttl,
'Only Mother Absent' = did_ttl_m_only,
#'Mother Absent' = #did_ttl_m,
'Only Father Absent' = did_ttl_f_only,
#'Father Absent' = #did_ttl_f,
'Both Parents Absent' = did_ttl_both) 
```

```{r}
table_ttl <- models_ttl %>% msummary(
  stars =  stars,
  fmt = 2,
  title = title_ttl,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  output = 'markdown')
table_ttl
```

z score
```{}
did_ttl_z <- feols(ttl_z_dif ~ migr + w1ttl_z | clsids + w2clsids, data = d_cl) 
did_ttl_z_m_only <- feols(ttl_z_dif ~ migr_mo_only + w1ttl_z | clsids + w2clsids, data = d_cl) 
did_ttl_z_m <- feols(ttl_z_dif ~ migr_mo + w1ttl_z | clsids + w2clsids, data = d_cl) 
did_ttl_z_f_only <- feols(ttl_z_dif ~ migr_fa_only + w1ttl_z | clsids + w2clsids, data = d_cl) 
did_ttl_z_f <- feols(ttl_z_dif ~ migr_fa + w1ttl_z | clsids + w2clsids, data = d_cl) 
did_ttl_z_both <- feols(ttl_z_dif ~ migr_both + w1ttl_z | clsids + w2clsids, data = d_cl)
```

```{}
models_ttl_z <- list(
'Any Parent Absent' = did_ttl_z,
'Only Mother Absent' = did_ttl_z_m_only,
#'Mother Absent' = did_ttl_z_m,
'Only Father Absent' = did_ttl_z_f_only,
#'Father Absent' = did_ttl_z_f,
'Both Parents Absent' = did_ttl_z_both) 
table_ttl_z <- models_ttl_z %>% msummary(stars = c('*' = .05, '**' = .01, '***' = .001), 
                              fmt = 2,
                              coef_rename =  c("migr" = "Explanator", "migr_mo_only" = "Explanator", "migr_mo" = "Explanator", "migr_fa" = "Explanator", "migr_fa_only" = "Explanator", "migr_both" = "Explanator", "w1ttl_dif" = "Baseline difference"), output = 'markdown')
table_ttl_z
```

std
```{}
did_ttl_std <- feols(ttl_std_dif ~ migr + w1ttl_std | clsids + w2clsids, data = d_cl) 
did_ttl_std_m_only <- feols(ttl_std_dif ~ migr_mo_only + w1ttl_std | clsids + w2clsids, data = d_cl) 
did_ttl_std_m <- feols(ttl_std_dif ~ migr_mo + w1ttl_std | clsids + w2clsids, data = d_cl) 
did_ttl_std_f_only <- feols(ttl_std_dif ~ migr_fa_only + w1ttl_std | clsids + w2clsids, data = d_cl) 
did_ttl_std_f <- feols(ttl_std_dif ~ migr_fa + w1ttl_std | clsids + w2clsids, data = d_cl) 
did_ttl_std_both <- feols(ttl_std_dif ~ migr_both + w1ttl_std | clsids + w2clsids, data = d_cl)
```

```{}
models_ttl_std <- list(
'Any Parent Absent' = did_ttl_std,
'Only Mother Absent' = did_ttl_std_m_only,
#'Mother Absent' = did_ttl_std_m,
'Only Father Absent' = did_ttl_std_f_only,
#'Father Absent' = did_ttl_std_f,
'Both Parents Absent' = did_ttl_std_both) 
table_ttl_std <- models_ttl_std %>% msummary(stars = c('*' = .05, '**' = .01, '***' = .001), 
                              fmt = 2,
                              coef_rename =  c("migr" = "Explanator", "migr_mo_only" = "Explanator", "migr_mo" = "Explanator", "migr_fa" = "Explanator", "migr_fa_only" = "Explanator", "migr_both" = "Explanator", "w1ttl_dif" = "Baseline difference"), output = 'markdown')
table_ttl_std
```

## DID adjusted unrestricted
$$\Delta score_{i,s} = \alpha + \beta \cdot migr_{i,s} + \gamma \cdot FE_{s} + \lambda \cdot score_{i,s;base} + \theta \cdot X_{i,s} + \varepsilon_{i,s}$$

Regressing on cognitive test score (standardized)
```{r}
did_adj_cog <- feols(cog_dif ~ migr + cog3pl  + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book  + home.has.internet | clsids + w2clsids, data = d_cl) 
did_adj_cog_m_only <- feols(cog_dif ~ migr_mo_only + cog3pl  + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book  + home.has.internet | clsids + w2clsids, data = d_cl) 
#did_adj_cog_m <- feols(cog_dif ~ migr_mo + cog3pl  + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book  + home.has.internet | clsids + w2clsids, data = d_cl) 
did_adj_cog_f_only <- feols(cog_dif ~ migr_fa_only + cog3pl + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book  + home.has.internet | clsids + w2clsids, data = d_cl) 
#did_adj_cog_f <- feols(cog_dif ~ migr_fa + cog3pl  + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book  + home.has.internet | clsids + w2clsids, data = d_cl) 
did_adj_cog_both <- feols(cog_dif ~ migr_both + cog3pl  + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book  + home.has.internet | clsids + w2clsids, data = d_cl)
```

Results
```{r}
models_adj_cog <- list(
'Any Parent Absent' = did_adj_cog,
'Only Mother Absent' = did_adj_cog_m_only,
#'Mother Absent' = #did_adj_cog_m,
'Only Father Absent' = did_adj_cog_f_only,
#'Father Absent' = #did_adj_cog_f,
'Both Parents Absent' = did_adj_cog_both) 
```

```{r}
table_adj_cog <- models_adj_cog %>% msummary(
  stars =  stars,
  fmt = 2,
  title = title_ttl,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  output = 'markdown'
  )
table_adj_cog
```

Regressing on total exam score
```{r}
did_adj_ttl <- feols(ttl_dif ~ migr + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet | clsids + w2clsids, data = d_cl) 
did_adj_ttl_m_only <- feols(ttl_dif ~ migr_mo_only + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet | clsids + w2clsids, data = d_cl) 
#did_adj_ttl_m <- feols(ttl_dif ~ migr_mo + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet | clsids + w2clsids, data = d_cl) 
#did_adj_ttl_f <- feols(ttl_dif ~ migr_fa + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet | clsids + w2clsids, data = d_cl) 
did_adj_ttl_f_only <- feols(ttl_dif ~ migr_fa_only + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet | clsids + w2clsids, data = d_cl) 
did_adj_ttl_both <- feols(ttl_dif ~ migr_both + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet | clsids + w2clsids, data = d_cl)
```

Results
```{r}
models_adj_ttl <- list(
'Any Parent Absent' = did_adj_ttl,
'Only Mother Absent' = did_adj_ttl_m_only,
#'Mother Absent' = #did_adj_ttl_m,
'Only Father Absent' = did_adj_ttl_f_only,
#'Father Absent' = #did_adj_ttl_f,
'Both Parents Absent' = did_adj_ttl_both)
```

```{r}
table_adj_ttl <- models_adj_ttl %>% msummary(
  stars =  stars,
  fmt = 2,
  title = title_ttl,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  output = 'markdown'
  )
table_adj_ttl
```

Regressing on total exam score
```{}
did_adj_ttl_z <- feols(ttl_z_dif ~ migr + w1ttl_z + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet | clsids + w2clsids, data = d_cl) 
did_adj_ttl_z_m_only <- feols(ttl_z_dif ~ migr_mo_only + w1ttl_z + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet | clsids + w2clsids, data = d_cl) 
did_adj_ttl_z_m <- feols(ttl_z_dif ~ migr_mo + w1ttl_z + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet | clsids + w2clsids, data = d_cl) 
did_adj_ttl_z_f <- feols(ttl_z_dif ~ migr_fa + w1ttl_z + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet | clsids + w2clsids, data = d_cl) 
did_adj_ttl_z_f_only <- feols(ttl_z_dif ~ migr_fa_only + w1ttl_z + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet | clsids + w2clsids, data = d_cl) 
did_adj_ttl_z_both <- feols(ttl_z_dif ~ migr_both + w1ttl_z + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet | clsids + w2clsids, data = d_cl)
```

Results
```{}
models_adj_ttl_z <- list(
'Any Parent Absent' = did_adj_ttl_z,
'Only Mother Absent' = did_adj_ttl_z_m_only,
#'Mother Absent' = did_adj_ttl_z_m,
'Only Father Absent' = did_adj_ttl_z_f_only,
#'Father Absent' = did_adj_ttl_z_f,
'Both Parents Absent' = did_adj_ttl_z_both) 
table_adj_ttl_z <- models_adj_ttl_z %>% msummary(stars = c('*' = .05, '**' = .01, '***' = .001), fmt = 2, coef_rename =  c("migr" = "Explanator", "migr_mo_only" = "Explanator", "migr_mo" = "Explanator", "migr_fa" = "Explanator", "migr_fa_only" = "Explanator", "migr_both" = "Explanator", "w1ttl_dif" = "Baseline difference"), output = 'markdown')
table_adj_ttl_z
```

Regressing on total exam score
```{}
did_adj_ttl_dm <- feols(ttl_dm_dif ~ migr + w1ttl_100_dm + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet | clsids + w2clsids, data = d_cl) 
did_adj_ttl_dm_m_only <- feols(ttl_dm_dif ~ migr_mo_only + w1ttl_100_dm + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet | clsids + w2clsids, data = d_cl) 
did_adj_ttl_dm_m <- feols(ttl_dm_dif ~ migr_mo + w1ttl_100_dm + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet | clsids + w2clsids, data = d_cl) 
did_adj_ttl_dm_f <- feols(ttl_dm_dif ~ migr_fa + w1ttl_100_dm + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet | clsids + w2clsids, data = d_cl) 
did_adj_ttl_dm_f_only <- feols(ttl_dm_dif ~ migr_fa_only + w1ttl_100_dm + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet | clsids + w2clsids, data = d_cl) 
did_adj_ttl_dm_both <- feols(ttl_dm_dif ~ migr_both + w1ttl_100_dm + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet | clsids + w2clsids, data = d_cl)
```

Results
```{}
models_adj_ttl_dm <- list(
'Any Parent Absent' = did_adj_ttl_dm,
'Only Mother Absent' = did_adj_ttl_dm_m_only,
#'Mother Absent' = did_adj_ttl_dm_m,
'Only Father Absent' = did_adj_ttl_dm_f_only,
#'Father Absent' = did_adj_ttl_dm_f,
'Both Parents Absent' = did_adj_ttl_dm_both) 
table_adj_ttl_dm <- models_adj_ttl_dm %>% msummary(stars = c('*' = .05, '**' = .01, '***' = .001), fmt = 2, coef_rename =  c("migr" = "Explanator", "migr_mo_only" = "Explanator", "migr_mo" = "Explanator", "migr_fa" = "Explanator", "migr_fa_only" = "Explanator", "migr_both" = "Explanator", "w1ttl_dif" = "Baseline difference"), output = 'markdown')
table_adj_ttl_dm
```

## Matching
```{r}
m.out <- matchit(
  migr ~ cog3pl + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.is.ethnic.minority + child.educational.aspiration + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet, 
  data = d_cl, 
  replace = T,
  ratio = 3,
  method = "nearest", 
  distance = "glm",
  exact = migr ~ ctyids
  )
fig_msmd_any <- summary(m.out) %>% plot(xlim = c(0, 0.55), position = 'right')
```

```{r}
dm <- match.data(m.out)
```

```{r}
didm_cog <- feols(cog_dif ~ migr + cog3pl | clsids + w2clsids, data = dm, weights = dm$weights)

didm_ttl <- feols(ttl_dif ~ migr + w1ttl_100 | clsids + w2clsids, data = dm, weights = dm$weights)

didm_adj_cog <- feols(cog_dif ~ migr + cog3pl + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet | clsids + w2clsids, data = dm, weights = dm$weights)

didm_adj_ttl <- feols(ttl_dif ~ migr + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet  | clsids + w2clsids, data = dm, weights = dm$weights)

```

```{r}
d_mo_only <- d_cl %>% drop_na(migr_mo_only)
m.out_mo_only <- 
  matchit(migr_mo_only ~ cog3pl + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.is.ethnic.minority + child.educational.aspiration + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet, 
  data = d_mo_only, 
  replace = T,
  method = "nearest",
  distance = "glm",
  ratio = 3,
  exact = migr_mo_only ~ ctyids
  )
fig_msmd_mo_only <- summary(m.out_mo_only) %>% plot(xlim = c(0, 0.55), position = 'right')
```

```{r}
dm_mo_only <- match.data(m.out_mo_only)
```

```{r}
didm_cog_mo_only <- feols(cog_dif ~ migr_mo_only + cog3pl | clsids + w2clsids, data = dm_mo_only, weights = dm_mo_only$weights)

didm_ttl_mo_only <- feols(ttl_dif ~ migr_mo_only + w1ttl_100 | clsids + w2clsids, data = dm_mo_only, weights = dm_mo_only$weights)

didm_adj_cog_mo_only <- feols(cog_dif ~ migr_mo_only + cog3pl + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet | clsids + w2clsids, data = dm_mo_only, weights = dm_mo_only$weights)

didm_adj_ttl_mo_only <- feols(ttl_dif ~ migr_mo_only + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet  | clsids + w2clsids, data = dm_mo_only, weights = dm_mo_only$weights)

```

```{r}
d_fa_only <- d_cl %>% drop_na(migr_fa_only)
m.out_fa_only <- matchit(
  migr_fa_only ~ cog3pl + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.is.ethnic.minority + child.educational.aspiration + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet,
  data = d_fa_only,
  replace = T,
  method = "nearest",
  distance = "glm",
  ratio = 3,
  exact = migr_fa_only ~ ctyids
  )
fig_msmd_fa_only <- summary(m.out_fa_only) %>% plot(xlim = c(0, 0.55), position = 'right')
```

```{r}
dm_fa_only <- match.data(m.out_fa_only)
```

```{r}
didm_cog_fa_only <- feols(cog_dif ~ migr_fa_only + cog3pl | clsids + w2clsids, data = dm_fa_only, weights = dm_fa_only$weights)

didm_ttl_fa_only <- feols(ttl_dif ~ migr_fa_only + w1ttl_100 | clsids + w2clsids, data = dm_fa_only, weights = dm_fa_only$weights)

didm_adj_cog_fa_only <- feols(cog_dif ~ migr_fa_only + cog3pl + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet | clsids + w2clsids, data = dm_fa_only, weights = dm_fa_only$weights)

didm_adj_ttl_fa_only <- feols(ttl_dif ~ migr_fa_only + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet  | clsids + w2clsids, data = dm_fa_only, weights = dm_fa_only$weights)

```

```{r}
d_both <- d_cl %>% drop_na(migr_both) 
m.out_both <- matchit(
  migr_both ~ cog3pl + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.is.ethnic.minority + child.educational.aspiration + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level + home.economic.resource + home.extracurricular.book + home.has.internet, 
  data = d_both, 
  method = "nearest",
  distance = "glm",
  replace = T,
  ratio = 3,
  exact = migr_both ~ ctyids
  )
fig_msmd_both <- plot(summary(m.out_both), xlim = c(0, 0.55), position = 'right')
```

```{r}
dm_both <- match.data(m.out_both)
```

```{r}
didm_cog_both <- feols(cog_dif ~ migr_both + cog3pl | clsids + w2clsids, data = dm_both, weights = dm_both$weights)

didm_ttl_both <- feols(ttl_dif ~ migr_both + w1ttl_100 | clsids + w2clsids, data = dm_both, weights = dm_both$weights)

didm_adj_cog_both <- feols(cog_dif ~ migr_both + cog3pl + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet | clsids + w2clsids, data = dm_both, weights = dm_both$weights)

didm_adj_ttl_both <- feols(ttl_dif ~ migr_both + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet  | clsids + w2clsids, data = dm_both, weights = dm_both$weights)

```

```{}
d_mo <- d_cl %>% drop_na(migr_mo)
m.out_mo <- matchit(
  migr_mo ~ cog3pl + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet, data = d_mo, 
  method = "nearest",
  replace = T,
  ratio = 3,
  exact = migr_mo ~ ctyids
  )
summary(m.out_mo, un = FALSE)
```

```{}
dm_mo <- match.data(m.out_mo)
```

```{}
feols(cog_dif ~ migr_mo + cog3pl| clsids + w2clsids, data = dm_mo, weights = dm_mo$weights)

feols(ttl_dif ~ migr_mo + w1ttl_100  | clsids + w2clsids, data = dm_mo, weights = dm_mo$weights)

feols(cog_dif ~ migr_mo + cog3pl + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet | clsids + w2clsids, data = dm_mo, weights = dm_mo$weights)

feols(ttl_dif ~ migr_mo + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet  | clsids + w2clsids, data = dm_mo, weights = dm_mo$weights)

```

```{}
d_fa <- d_cl %>% drop_na(migr_fa)
m.out_fa <- matchit(
  migr_fa ~ cog3pl + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet, data = d_fa,
  method = "nearest",
  replace = T,
  ratio = 3,
  exact = migr_fa_only ~ ctyids
  )
summary(m.out_fa, un = FALSE)
```

```{}
dm_fa <- match.data(m.out_fa)
```

```{}
feols(cog_dif ~ migr_fa + cog3pl | clsids + w2clsids, data = dm_fa, weights = dm_fa$weights)

feols(ttl_dif ~ migr_fa + w1ttl_100 | clsids + w2clsids, data = dm_fa, weights = dm_fa$weights)

feols(cog_dif ~ migr_fa + cog3pl + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet | clsids + w2clsids, data = dm_fa, weights = dm_fa$weights)

feols(ttl_dif ~ migr_fa + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet  | clsids + w2clsids, data = dm_fa, weights = dm_fa$weights)

```

```{r}
title_cog_matching <- 'Parental migration’s effect on children’s cognitive abilities, estimated with matching'
title_ttl_matching <- 'Parental migration’s effect on children’s academic abilities, estimated with matching'
note_matching <- "3:1 nearest neighbor propensity score matching with replacement; exact matching by county"
```

```{r}
models_m_cog <- list(
'Any Parent Absent' = didm_cog,
'Only Mother Absent' = didm_cog_mo_only,
##'Mother Absent' = #did_adj_cog_m,
'Only Father Absent' = didm_cog_fa_only,
##'Father Absent' = #did_adj_cog_f,
'Both Parents Absent' = didm_cog_both) 
```

```{r}
table_m_cog <- models_m_cog %>% msummary(
  stars =  stars,
  fmt = 2,
  title = title_cog_matching,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  notes = note_matching,
  output = 'markdown'
  )
table_m_cog
```

```{r}
models_m_ttl <- list(
'Any Parent Absent' = didm_ttl,
'Only Mother Absent' = didm_ttl_mo_only,
##'Mother Absent' = #did_adj_ttl_m,
'Only Father Absent' = didm_ttl_fa_only,
##'Father Absent' = #did_adj_ttl_f,
'Both Parents Absent' = didm_ttl_both) 
table_m_ttl <- models_m_ttl %>% msummary(
  stars =  stars,
  fmt = 2,
  title = title_ttl_matching,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  notes = note_matching,
  output = 'markdown'
  )
table_m_ttl
```

```{r}
models_m_adj_cog <- list(
'Any Parent Absent' = didm_adj_cog,
'Only Mother Absent' = didm_adj_cog_mo_only,
##'Mother Absent' = #did_adj_cog_m,
'Only Father Absent' = didm_adj_cog_fa_only,
##'Father Absent' = #did_adj_cog_f,
'Both Parents Absent' = didm_adj_cog_both) 
table_m_adj_cog <- models_m_adj_cog %>% msummary(
  stars =  stars,
  fmt = 2,
  title = title_ttl_matching,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  notes = note_matching,
  output = 'markdown'
  )
table_m_adj_cog
```

```{r}
models_m_adj_ttl <- list(
'Any Parent Absent' = didm_adj_ttl,
'Only Mother Absent' = didm_adj_ttl_mo_only,
##'Mother Absent' = #did_adj_ttl_m,
'Only Father Absent' = didm_adj_ttl_fa_only,
##'Father Absent' = #did_adj_ttl_f,
'Both Parents Absent' = didm_adj_ttl_both) 
table_m_adj_ttl <- models_m_adj_ttl %>% msummary(
  stars =  stars,
  fmt = 2,
  title = title_ttl_matching,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  notes = note_matching,
  output = 'markdown'
  )
table_m_adj_ttl
```

```{r}
num.outlier <- d_cl%>%filter(w1mat_100>100|w1eng_100>100|w2mat_100>100|w2eng_100>100)%>%count
```

## Interaction
```{r}
d_cl1 <- d_cl %>% mutate(
  home.economic.resource.w2 = if_else(is.na(w2be23), w2a09, w2be23)
)
d_cl1 <- d_cl %>% mutate(
  fin.better = if_else(home.economic.resource.w2 > home.economic.resource,
                       1,
                       0)
  )

```

```{r}
feols(ttl_dif ~ migr*fin.better + w1ttl_100 | clsids + w2clsids, data = d_cl1) %>%
  msummary(stars = c('†' = .1, '*' = .05, '**' = .01, '***' = .001),
           output = 'markdown')
feols(cog_dif ~ migr*fin.better + cog3pl | clsids + w2clsids, data = d_cl1)  %>%
  msummary(stars = c('†' = .1, '*' = .05, '**' = .01, '***' = .001),
           output = 'markdown')
```

```{r}
feols(ttl_dif ~ migr*fin.better + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet  | clsids + w2clsids, data = d_cl1) %>%
    msummary(stars = c('†' = .1, '*' = .05, '**' = .01, '***' = .001),
           output = 'markdown')
feols(cog_dif ~ migr*fin.better + cog3pl + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet  | clsids + w2clsids, data = d_cl1) %>%
    msummary(stars = c('†' = .1, '*' = .05, '**' = .01, '***' = .001),
           output = 'markdown')
```

```{r}
d_cl1=d_cl1 %>% mutate(
  supervision.less = if_else((b2201 + b2202) > (w2a18 + w2a19), 1, 0)
)
```

```{r}
feols(ttl_dif ~ migr*supervision.less + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet  | clsids + w2clsids, data = d_cl1) %>%
    msummary(stars = c('†' = .1, '*' = .05, '**' = .01, '***' = .001),
           output = 'markdown')
feols(cog_dif ~ migr*supervision.less + cog3pl + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.educational.aspiration + child.is.ethnic.minority + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet  | clsids + w2clsids, data = d_cl1) %>%
    msummary(stars = c('†' = .1, '*' = .05, '**' = .01, '***' = .001),
           output = 'markdown')
```