### 2022-07-03 data/Gu-thesis-script.Rmd

---
title: 'parental migration and adolescents educational performance'
output: html_notebook
---

Load packages
```{r}
library(tidyverse) 
library(haven)
library(fixest)
library(modelsummary)
library(MatchIt)
library(vtable)
library(labelled)
library(cobalt) 
options(modelsummary_get="broom")
#library(gtsummary)
#library(DescTools)
#knitr::opts_chunk$set(modelsummary_get="broom")
```

## Data cleaning
Read CEPS data
```{r}
d_stu1 <- read_dta('cepsw1studentCN.dta') #wave 1 student
d_stu2 <- read_dta('cepsw2studentCN.dta') #wave 2 student
d_par1 <- read_dta('cepsw1parentCN.dta') #wave 1 parent
d_par2 <- read_dta('cepsw2parentCN.dta') #wave 2 parent
#d_sch1 <- read_dta('cepsw1principalCN.dta') #wave 1 school principal
#d_sch2 <- read_dta('cepsw2principalCN.dta')
```

Remove wave 1's observations of grade 9; they were not followed up in wave 2
```{r}
d_stu1 <- d_stu1 |> 
  subset(grade9==0) |> 
  select(!c(frame, subsample, sweight, grade9))
d_stu2 <- d_stu2 |> 
  select(!c(w2frame, w2subsample, w2sweight, w1w2sweight))
d_par1 <- d_par1 |> 
  subset(grade9==0) |> 
  select(!c(frame, subsample, grade9))
```

Assemble a panel dataset in wide format
```{r}
dw1 <- d_stu1 |> left_join(d_par1, by=c('ids', 'schids', 'clsids', 'ctyids', 'fall')) 
dw2 <- d_stu2 |> left_join(d_par2, by=c('ids', 'clsids', 'w2clsids', 'schids', 'ctyids', 'w2fall'))
d <- dw1 |> right_join(dw2, by=c('ids', 'clsids', 'schids', 'ctyids')) |> remove_labels()
```

```{r}
d <- d |> mutate(
  w1chn_100 = tr_chn/w2upchn*100, #raw exam scores at 100-point scale
  w1mat_100 = tr_mat/w2upmat*100,
  w1eng_100 = tr_eng/w2upeng*100,
  w2chn_100 = w2chn/w2upchn*100,
  w2mat_100 = w2mat/w2upmat*100,
  w2eng_100 = w2eng/w2upeng*100
)
```

Trimming outliers
```{r}
d <- d |> mutate(
  w1chn_100_trim = if_else(w1chn_100>100, 100, w1chn_100),
  w2chn_100_trim = if_else(w2chn_100>100, 100, w2chn_100),
  w1mat_100_trim = if_else(w1mat_100>100, 100, w1mat_100),
  w2mat_100_trim = if_else(w2mat_100>100, 100, w2mat_100),
  w1eng_100_trim = if_else(w1eng_100>100, 100, w1eng_100),
  w2eng_100_trim = if_else(w2eng_100>100, 100, w2eng_100)
)
```

```{r}
d <- d |> mutate(
  w1ttl_100 = w1chn_100_trim + w1mat_100_trim + w1eng_100_trim,
  w2ttl_100 = w2chn_100_trim + w2mat_100_trim + w2eng_100_trim,
  w1avg_100 = w1ttl_100/3,
  w2avg_100 = w2ttl_100/3,
  w1cog_100 = stcog*100/20,
  w2cog_100 = w2cogscore*100/35,
  ttl_dif = w2ttl_100 - w1ttl_100,
  avg_dif = w2avg_100 - w1avg_100,
  cog_dif = w2cog_100 - w1cog_100,
  #cog_std_dif = w2cog3pl - cog3pl
  )
```

```{r}
d_cl <- d |>
  drop_na(cog_dif, avg_dif)
```

within-class z score
```{}
d1 <- d1 |> group_by(schids) |>
  mutate(
    w1ttl_z = as.numeric(scale(w1ttl)),
    w2ttl_z = as.numeric(scale(w2ttl))
  ) |> ungroup()
```

```{}
d1 <- d1 |> group_by(schids) |>
  mutate(
    w1ttl_std = rescale(w1ttl),
    w2ttl_std = rescale(w2ttl)
    ) |> ungroup()
```

Detect outliers
```{}
d_outlier <- d |> filter(w1chn_100 > 100 | w1mat_100>100|w1eng_100>100|w2chn_100>100|w2mat_100>100|w2eng_100>100)
d_outlier1 <- d_outlier |> select(clsids, schids, tr_chn, tr_mat, tr_eng, w2chn, w2upchn, w2mat, w2upmat, w2eng, w2upeng)
d1 <- d |> anti_join(d_outlier, by=c('ids', 'clsids', 'schids', 'ctyids'))
```

Sample
```{r}
d_cl <- d |> 
  drop_na(cog_dif, avg_dif) |> 
  filter(w2a0801==1) |> #keep parents who are married (and alive)
  filter(be1201==1 & be1202==1) |> #keep both parents live with children at baseline
  filter(stmigrant==1) #keep local children 
```

## Variables

Control variables
```{r}
d_cl <- d_cl |> mutate(
         child.age = 2013 - a02a, #student's age at baseline
         child.is.girl = if_else(a01 == 2, 1, 0), #student is girl
         child.has.rural.hukou = sthktype, #student holds rural (agricultural) hukou
         child.in.boarding.school = stbrd, #student is boarding at school
         child.is.ethnic.minority = if_else(a03 != 1, 1, 0), #student is ethnic minority
         child.health = a17, #student's self-rated health
         child.had.skipped.grade = if_else(c03 == 0, 0, 1), #student had skipped grade
         child.had.repeated.grade = if_else(c04 == 0, 0, 1), #student had repeated grade
         #ch.illness = if_else(a10 == 1, 1, 0), #student had serious illness before primary school
         child.went.to.preschool = if_else(c01 == 1, 1, 0), #student had attended preschool
         child.number.of.sibling = if_else(stonly == 1, 0, stsib),
         #ch.par.rltn = if_else(stprrel == 2, 1, 0),
         #ch.tutoring = if_else(b1901 == 1|b1902 == 1|b1903 == 1|b1904 == 1, 1, 0),
         parent.has.white.collar.job = if_else(be08==1|be08==2|be08==3, 1, 0),
         #par.ccp = if_else(be06 == 1, 1, 0), #At least one parent is communist party member
         home.economic.resource = steco_5c, #self-rated household financial situation
         home.has.internet = if_else(b13 == 2, 1, 0), #student has internet access at home
         home.extracurricular.book = b12 #Amount of book at home, student self-rated 
) |> drop_na(
child.age, child.is.girl, child.has.rural.hukou, child.health, child.in.boarding.school, child.is.ethnic.minority, child.health, child.had.skipped.grade, child.had.repeated.grade, child.went.to.preschool, child.number.of.sibling, 
#par.ccp, 
parent.has.white.collar.job, home.has.internet, home.extracurricular.book, home.economic.resource)
```

```{}
d_cl$stsib <- d_cl$stsib |> replace_na(0) |> as.numeric()
d_cl <- d_cl |> mutate(child.number.of.sibling = if_else(stonly == 1, 0, stsib)) |> #number of siblings
  drop_na(child.number.of.sibling)
```

Best-educated parent's years of education
```{r}
d_cl <- d_cl |> mutate(parent.education.level = NA)
d_cl$parent.education.level[d_cl$stprhedu == 1|d_cl$stprhedu == 2] <- 0
d_cl$parent.education.level[d_cl$stprhedu == 3] <- 1
d_cl$parent.education.level[d_cl$stprhedu == 4|d_cl$stprhedu == 5|d_cl$stprhedu == 6] <- 2
d_cl$parent.education.level[d_cl$stprhedu == 7] <- 3
d_cl$parent.education.level[d_cl$stprhedu == 8] <- 4
d_cl$parent.education.level[d_cl$stprhedu == 9] <- 5
```

```{r}
d_cl <- d_cl |> mutate(mother.education.level = NA)
d_cl$mother.education.level[d_cl$stmedu == 1|d_cl$stmedu == 2] <- 0
d_cl$mother.education.level[d_cl$stmedu == 3] <- 1
d_cl$mother.education.level[d_cl$stmedu == 4|d_cl$stmedu == 5|d_cl$stmedu == 6] <- 2
d_cl$mother.education.level[d_cl$stmedu == 7] <- 3
d_cl$mother.education.level[d_cl$stmedu == 8] <- 4
d_cl$mother.education.level[d_cl$stmedu == 9] <- 5
```

```{r}
d_cl <- d_cl |> mutate(father.education.level = NA)
d_cl$father.education.level[d_cl$stfedu == 1|d_cl$stfedu == 2] <- 0
d_cl$father.education.level[d_cl$stfedu == 3] <- 1
d_cl$father.education.level[d_cl$stfedu == 4|d_cl$stfedu == 5|d_cl$stfedu == 6] <- 2
d_cl$father.education.level[d_cl$stfedu == 7] <- 3
d_cl$father.education.level[d_cl$stfedu == 8] <- 4
d_cl$father.education.level[d_cl$stfedu == 9] <- 5
```

```{r}
d_cl <- d_cl |> mutate(parent.expectation = NA)
d_cl$parent.expectation[d_cl$ba18 == 1|d_cl$ba18 == 2] <- 1
d_cl$parent.expectation[d_cl$ba18 == 3|d_cl$ba18 == 4|d_cl$ba18 == 5] <- 2
d_cl$parent.expectation[d_cl$ba18 == 6] <- 3
d_cl$parent.expectation[d_cl$ba18 == 7] <- 4
d_cl$parent.expectation[d_cl$ba18 == 8|d_cl$ba18 == 9] <- 5
```

```{r}
d_cl <- d_cl |> mutate(child.education.aspiration = NA)
d_cl$child.education.aspiration[d_cl$c22 == 1|d_cl$c22 == 2|d_cl$c22 == 10] <- 1
d_cl$child.education.aspiration[d_cl$c22 == 3|d_cl$c22 == 4|d_cl$c22 == 5] <- 2
d_cl$child.education.aspiration[d_cl$c22 == 6] <- 3
d_cl$child.education.aspiration[d_cl$c22 == 7] <- 4
d_cl$child.education.aspiration[d_cl$c22 == 8|d_cl$c22 == 9] <- 5
```

```{r}
d_cl <- d_cl |>  drop_na(
  parent.education.level, 
  parent.expectation, 
  child.education.aspiration)
```

## Treatment groups
At least one parent migrated (Control group as no parent migrated)
```{r}
d_cl <- d_cl |> mutate(
  migr = if_else((w2ba0201==0 | w2ba0202==0), 1, 0), #Any parent became Migrated from household between baseline and endline survey
  migr_fa_only = NA,
  #migr_fa = NA,
  migr_mo_only = NA,
  #migr_mo = NA,
  migr_both = NA
  ) |> drop_na(migr)
```

Only mother migrated
```{r}
d_cl$migr_mo_only[d_cl$migr == 0] <- 0
d_cl$migr_mo_only[d_cl$w2ba0201==1 & d_cl$w2ba0202==0] <- 1 
```

Mother migrated
```{}
d_cl$migr_mo[d_cl$migr == 0] <- 0
d_cl$migr_mo[d_cl$w2ba0202==0] <- 1
```

Only father migrated
```{r}
d_cl$migr_fa_only[d_cl$migr == 0] <- 0
d_cl$migr_fa_only[d_cl$w2ba0201==0 & d_cl$w2ba0202==1] <- 1 
```

Father migrated
```{}
d_cl$migr_fa[d_cl$migr == 0] <- 0
d_cl$migr_fa[d_cl$w2ba0201==0] <- 1
```

Both parents migrated
```{r}
d_cl$migr_both[d_cl$migr == 0] <- 0
d_cl$migr_both[d_cl$w2ba0201==0 & d_cl$w2ba0202==0] <- 1
```

```{}
df <- d_cl |> select(
  migr, migr_fa_only, migr_fa, migr_mo_only, migr_mo, migr_both,
  cog3pl, w1ttl_100, cog_dif, ttl_dif,
  child.is.girl, child.has.rural.hukou, child.health, child.in.boarding.school, child.is.ethnic.minority, child.health, child.had.skipped.grade, child.had.repeated.grade, child.went.to.preschool, child.education.aspiration, 
  par.edu, par.ccp, par.expt.univ, 
  home.has.internet, home.extracurricular.book, home.economic.resource
)
```

## Descriptive statistics
```{r}
d_migr <- d_cl |>
  select(
  migr, migr_mo_only, migr_fa_only, migr_both,
  w1cog_100, w2cog_100, w1avg_100, w2avg_100, cog_dif, avg_dif,
  child.age, child.is.girl, child.has.rural.hukou, child.in.boarding.school, child.is.ethnic.minority, child.had.skipped.grade, child.had.repeated.grade, child.went.to.preschool, parent.has.white.collar.job, home.has.internet,
  child.health, child.education.aspiration, 
  mother.education.level, father.education.level, parent.expectation, 
  home.extracurricular.book, home.economic.resource) 
```

```{r}
d_smr <- d_migr |> 
  mutate(`Parental Migration` = if_else((migr==1), 'Yes', 'No')) |> 
  mutate(`Parental Migration Type` = 0) 
d_smr$`Parental Migration Type`[d_smr$migr==0] <- 'Both Parents At Home'
d_smr$`Parental Migration Type`[d_smr$migr_mo_only==1] <- 'Only Mother'
d_smr$`Parental Migration Type`[d_smr$migr_fa_only==1] <- 'Only Father'
d_smr$`Parental Migration Type`[d_smr$migr_both==1] <- 'Both Parents'
d_smr <- d_smr |> select(!migr & !migr_mo_only & !migr_fa_only & !migr_both)
```

```{}
datasummary_balance(
  ~migr,
  data = d_migr,
  fmt = 2)
```

```{r}
d_dpvar1 <- d_migr |>
  select(migr, 
         cog_dif, w1cog_100, w2cog_100, avg_dif, w1avg_100, w2avg_100) |>
  set_variable_labels(
    migr = 'Parental migration',
    w1cog_100 ='Wave 1 score',
    w2cog_100 = 'Wave 2 score',
    w1avg_100 ='Wave 1 score',
    w2avg_100 = 'Wave 2 score',
    cog_dif = 'Cognitive test score difference',
    avg_dif = 'Academic exam score difference'
  )
```

```{r}
tab_dpvar <- 
  st(
  data = d_dpvar1,
  summ = c('mean(x)', 'sd(x)'),
  digits = 2, 
  group = 'migr',
  group.test = list(
    star.cutoffs = c(0.001, 0.01, 0.05),
    format = 'p={pval}{stars}'),
  labels = T,
  out = 'return') 
tab_dpvar <- tab_dpvar[-c(1),]
tab_dpvar[tab_dpvar =='p=&lt0.001***']<-'p<0.001***'
tab_dpvar 
```

```{}
d_dpvar <- as.data.frame(d_dpvar)
table_dependent_variable <- datasummary(
  All(d_dpvar) ~ Mean + SD + Min + Max,
  data = d_dpvar,
  output = 'latex'
  ) 
table_dependent_variable
```

```{r}
d_idpvar <- d_migr |> 
  select(migr, migr_mo_only, migr_fa_only, migr_both) |> 
  mutate(`Parental Migration` = if_else((migr==1), 'Yes', 'No')) |> 
  mutate(`Parental Migration Type` = 0) 
d_idpvar$`Parental Migration Type`[d_idpvar$migr==0] <- 'Both Parents At Home'
d_idpvar$`Parental Migration Type`[d_idpvar$migr_mo_only==1] <- 'Only Mother'
d_idpvar$`Parental Migration Type`[d_idpvar$migr_fa_only==1] <- 'Only Father'
d_idpvar$`Parental Migration Type`[d_idpvar$migr_both==1] <- 'Both Parents'
d_idpvar <- d_idpvar |> select(!migr & !migr_mo_only & !migr_fa_only & !migr_both)
```

```{r}
tab_idpvar <- st(
  data = d_idpvar,
  out = 'return'
)
tab_idpvar[tab_idpvar == 4062] <- ''
tab_idpvar
```

```{r}
table_independent_variable <- datasummary_skim(
    data=d_idpvar,
    type = 'categorical',
    title = 'Independent variables',
    output = 'markdown')
table_independent_variable
```

```{r}
d_ctvar_dit <- d_migr |>
  select(migr, child.is.girl, child.has.rural.hukou, child.in.boarding.school, child.is.ethnic.minority, child.had.skipped.grade, child.had.repeated.grade, child.went.to.preschool, parent.has.white.collar.job, home.has.internet) |> 
  mutate_all(as_factor)
```

```{r}
d_ctvar_cnt <- d_migr |>
  select(migr, child.age, child.health, child.education.aspiration, 
  mother.education.level, father.education.level, parent.expectation, 
  home.extracurricular.book, home.economic.resource)
```

```{r}
d_ctvar_label <- d_ctvar_cnt|>
  select(!(child.age:migr)) |> 
  add_value_labels(
    father.education.level = c(
     '0 = Primary school and below' = 0,
     '1 = Junior high school' = 1 ,
     '2 = Senior high school' = 2,
     '3 = Associate college' = 3,
     '4 = University' = 4,
     '5 = Postgraduate' = 5),
    mother.education.level = c(
     '0 = Primary school and below' = 0,
     '1 = Junior high school' = 1 ,
     '2 = Senior high school' = 2,
     '3 = Associate college' = 3,
     '4 = University' = 4,
     '5 = Postgraduate' = 5),
    child.education.aspiration = c(
     '1 = Junior high school' = 1 ,
     '2 = Senior high school' = 2,
     '3 = Associate college' = 3,
     '4 = University' = 4,
     '5 = Postgraduate' = 5),
    child.education.aspiration = c(
     '1 = Junior high school' = 1 ,
     '2 = Senior high school' = 2,
     '3 = Associate college' = 3,
     '4 = University' = 4,
     '5 = Postgraduate' = 5), 
    parent.expectation = c(
     '1 = Junior high school' = 1 ,
     '2 = Senior high school' = 2,
     '3 = Associate college' = 3,
     '4 = University' = 4,
     '5 = Postgraduate' = 5),
    home.economic.resource = c(
      '1 = Very poor' = 1,
      '2 = Somewhat poor' = 2,
      '3 = Moderate' = 3,
      '4 = Somewhat rich' = 4,
      '5 = Very rich' = 5),
    child.health = c(
      '1 = Very poor' = 1,
      '2 = Somewhat poor' = 2,
      '3 = Moderate' = 3,
      '4 = Good' = 4,
      '5 = Very good' = 5),
    home.extracurricular.book = c(
      '1 = Very few' = 1,
      '2 = Not many' = 2,
      '3 = Some' = 3,
      '4 = Quite a few' = 4,
      '5 = A great number' = 5),
    ) |>
  as_factor()
```

```{}
test.table <- d_ctvar |>
  tbl_summary(
    by = migr,
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
    digits = all_continuous() ~ 2,
    ) 
```

```{r}
table_variable <- datasummary_skim(
  data = d_ctvar_label,
  type = 'categorical',
  title = 'Measurement of categorical variables',
  output = 'latex'
)
#table_variable
```

```{r}
tab_var <- st(
  data = d_ctvar_label,
  out = 'return'
)
tab_var[tab_var == 4062] <- ''
tab_var
```

```{r}
tab_ctvar_dit <- 
  st(
  data = d_ctvar_dit,
  digits = 2,
  group = 'migr',
  group.test = list(star.cutoffs = c(0.001, 0.01, 0.05)),
  out = "return"
  )
tab_ctvar_dit <- tab_ctvar_dit[-c(1),]
tab_ctvar_dit[tab_ctvar_dit == 3422|tab_ctvar_dit == 640] <- ''
tab_ctvar_dit 
```

```{r}
tab_ctvar_cnt <- d_ctvar_cnt |> st(
  digits = 2, 
  summ = c('mean(x)', 'sd(x)'),
  group = 'migr',
  group.test = list(star.cutoffs = c(0.001, 0.01, 0.05)),
  out = "return")
tab_ctvar_cnt <- tab_ctvar_cnt[-c(1),]
tab_ctvar_cnt
```

```{}
table_control_variable <- datasummary_skim(
  data = d_ctvar,
  type = 'categorical',
  output = 'latex'
)
table_control_variable
```

t-test
```{}
ttest_cog <- t.test(cog_dif ~ migr, data = d_cl)
ttest_ttl <- t.test(avg_dif ~ migr, data = d_cl)
ttest_cog
ttest_ttl
```

```{r}
set.seed(1)
fig_cog <- d_smr |>
  ggplot(aes(x=`Parental Migration`, y=cog_dif, fill=`Parental Migration`)) + 
    geom_violin()+
    xlab('Parental Migration') +
    theme(legend.position="none") +
    ylab('Cognitive Test Score Difference') +
    stat_summary(
      fun = "mean",
      geom = "crossbar", 
      size = 0.4,
      width = 0.5) +
  scale_fill_manual(values = alpha(c("green", "red"), .4))
fig_cog
```

```{r}
set.seed(1)
fig_ttl <-  d_smr |>
  ggplot(aes(x=`Parental Migration`, y=avg_dif, fill=`Parental Migration`)) + 
    geom_violin()+
    xlab('Parental Migration') +
    theme(legend.position="none") +
    ylab('Academic Exam Score Difference') +
    stat_summary(
      fun = "mean",
      geom = "crossbar", 
      size = 0.4,
      width = 0.5) +
  scale_fill_manual(values = alpha(c("green", "red"), .4))
fig_ttl
```

```{r}
set.seed(1)
fig_cog_by_migration_type <- d_smr |>
  ggplot(aes(x=`Parental Migration Type`, y=cog_dif, fill = `Parental Migration Type`)) + 
    geom_violin()+
    xlab('Parental Migration Type') +
    theme(legend.position="none") +
    ylab('Cognitive Test Score Difference') +
    stat_summary(
      fun = "mean",
      geom = "crossbar", 
      width = 0.5)+
  scale_fill_manual(values = alpha(c("green", "tomato1", "tomato2", "tomato3"), .4))
fig_cog_by_migration_type
```

```{r}
set.seed(1)
fig_ttl_by_migration_type <- d_smr |>
  ggplot(aes(x=`Parental Migration Type`, y=avg_dif, fill = `Parental Migration Type`)) + 
    geom_violin()+
    xlab('Parental Migration Type') +
    theme(legend.position="none") +
    ylab('Academic Exam Score Difference') +
    stat_summary(
      fun = "mean",
      geom = "crossbar", 
      width = 0.5) +
  scale_fill_manual(values = alpha(c("green", "tomato1", "tomato2", "tomato3"), .4))
fig_ttl_by_migration_type
```

## DID unrestricted
Regressing on cognitive test score (standardized)
```{r}
did_cog <- feols(cog_dif ~ migr + w1cog_100 | schids, data = d_cl) 
did_cog_m_only <- feols(cog_dif ~ migr_mo_only + w1cog_100 | schids, data = d_cl) 
#did_cog_m <- feols(cog_dif ~ migr_mo + w1cog_100 | schids, data = d_cl) 
did_cog_f_only <- feols(cog_dif ~ migr_fa_only + w1cog_100 | schids, data = d_cl) 
#did_cog_f <- feols(cog_dif ~ migr_fa + w1cog_100 | schids, data = d_cl) 
did_cog_both <- feols(cog_dif ~ migr_both + w1cog_100 | schids, data = d_cl)
```

Results
```{r}
models_cog <- list(
'Any Parent' = did_cog,
'Only Mother' = did_cog_m_only,
#'Mother' = #did_cog_m,
'Only Father' = did_cog_f_only,
#'Father' = #did_cog_f,
'Both Parents' = did_cog_both) 
```

```{r}
stars <- c('†' = .1, '*' = .05, '**' = .01, '***' = .001)
coef_rename <- c("migr" = "Migration", "migr_mo_only" = "Migration", "migr_mo" = "Migration", "migr_fa" = "Migration", "migr_fa_only" = "Migration", "migr_both" = "Migration", "w1cog_100" = "Baseline.score", "w1avg_100" = "Baseline.score")
title_cog <- 'Parental migration’s effect on children’s cognitive test score'
title_ttl <- 'Parental migration’s effect on children’s academic exam score'
gof_omit <- 'R2 Within|R2 Pseudo|AIC|BIC|Log.Lik.|Std.Errors'
note_unadj <- 'No control variables included in these models'
```

```{r}
table_cog <- models_cog |> msummary(
  stars =  stars,
  fmt = 2,
  title = title_cog,
  gof_omit = gof_omit,
  coef_rename =  coef_rename,
  notes = note_unadj,
  output = 'latex')
#table_cog
```

raw score
```{r}
did_ttl <- feols(avg_dif ~ migr + w1avg_100 | schids, data = d_cl) 
did_ttl_m_only <- feols(avg_dif ~ migr_mo_only + w1avg_100 | schids, data = d_cl) 
#did_ttl_m <- feols(avg_dif ~ migr_mo + w1avg_100 | schids, data = d_cl) 
did_ttl_f_only <- feols(avg_dif ~ migr_fa_only + w1avg_100 | schids, data = d_cl) 
#did_ttl_f <- feols(avg_dif ~ migr_fa + w1avg_100 | schids, data = d_cl) 
did_ttl_both <- feols(avg_dif ~ migr_both + w1avg_100 | schids, data = d_cl)
```

```{r}
models_ttl <- list(
'Any Parent' = did_ttl,
'Only Mother' = did_ttl_m_only,
#'Mother' = #did_ttl_m,
'Only Father' = did_ttl_f_only,
#'Father' = #did_ttl_f,
'Both Parents' = did_ttl_both) 
```

```{r}
table_ttl <- models_ttl |> msummary(
  stars =  stars,
  fmt = 2,
  title = title_ttl,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  notes = note_unadj,
  output = 'latex')
#table_ttl
```

z score
```{}
did_ttl_z <- feols(ttl_z_dif ~ migr + w1ttl_z | schids, data = d_cl) 
did_ttl_z_m_only <- feols(ttl_z_dif ~ migr_mo_only + w1ttl_z | schids, data = d_cl) 
did_ttl_z_m <- feols(ttl_z_dif ~ migr_mo + w1ttl_z | schids, data = d_cl) 
did_ttl_z_f_only <- feols(ttl_z_dif ~ migr_fa_only + w1ttl_z | schids, data = d_cl) 
did_ttl_z_f <- feols(ttl_z_dif ~ migr_fa + w1ttl_z | schids, data = d_cl) 
did_ttl_z_both <- feols(ttl_z_dif ~ migr_both + w1ttl_z | schids, data = d_cl)
```

```{}
models_ttl_z <- list(
'Any Parent' = did_ttl_z,
'Only Mother' = did_ttl_z_m_only,
#'Mother' = did_ttl_z_m,
'Only Father' = did_ttl_z_f_only,
#'Father' = did_ttl_z_f,
'Both Parents' = did_ttl_z_both) 
table_ttl_z <- models_ttl_z |> msummary(stars = c('*' = .05, '**' = .01, '***' = .001), 
                              fmt = 2,
                              coef_rename =  c("migr" = "Migration", "migr_mo_only" = "Migration", "migr_mo" = "Migration", "migr_fa" = "Migration", "migr_fa_only" = "Migration", "migr_both" = "Migration", "w1ttl_dif" = "Baseline difference"), output = 'latex')
table_ttl_z
```

std
```{}
did_ttl_std <- feols(ttl_std_dif ~ migr + w1ttl_std | schids, data = d_cl) 
did_ttl_std_m_only <- feols(ttl_std_dif ~ migr_mo_only + w1ttl_std | schids, data = d_cl) 
did_ttl_std_m <- feols(ttl_std_dif ~ migr_mo + w1ttl_std | schids, data = d_cl) 
did_ttl_std_f_only <- feols(ttl_std_dif ~ migr_fa_only + w1ttl_std | schids, data = d_cl) 
did_ttl_std_f <- feols(ttl_std_dif ~ migr_fa + w1ttl_std | schids, data = d_cl) 
did_ttl_std_both <- feols(ttl_std_dif ~ migr_both + w1ttl_std | schids, data = d_cl)
```

```{}
models_ttl_std <- list(
'Any Parent' = did_ttl_std,
'Only Mother' = did_ttl_std_m_only,
#'Mother' = did_ttl_std_m,
'Only Father' = did_ttl_std_f_only,
#'Father' = did_ttl_std_f,
'Both Parents' = did_ttl_std_both) 
table_ttl_std <- models_ttl_std |> msummary(stars = c('*' = .05, '**' = .01, '***' = .001), 
                              fmt = 2,
                              coef_rename =  c("migr" = "Migration", "migr_mo_only" = "Migration", "migr_mo" = "Migration", "migr_fa" = "Migration", "migr_fa_only" = "Migration", "migr_both" = "Migration", "w1ttl_dif" = "Baseline difference"), output = 'latex')
table_ttl_std
```

## DID adjusted unrestricted
$$\Delta score_{i,s} = \alpha + \beta \cdot migr_{i,s} + \gamma \cdot FE_{s} + \lambda \cdot score_{i,s;base} + \theta \cdot X_{i,s} + \varepsilon_{i,s}$$

Regressing on cognitive test score 
```{r}
did_adj_cog <- feols(cog_dif ~ migr + w1cog_100  + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book  + home.has.internet | schids, data = d_cl) 
did_adj_cog_m_only <- feols(cog_dif ~ migr_mo_only + w1cog_100  + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book  + home.has.internet | schids, data = d_cl) 
#did_adj_cog_m <- feols(cog_dif ~ migr_mo + w1cog_100  + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book  + home.has.internet | schids, data = d_cl) 
did_adj_cog_f_only <- feols(cog_dif ~ migr_fa_only + w1cog_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book  + home.has.internet | schids, data = d_cl) 
#did_adj_cog_f <- feols(cog_dif ~ migr_fa + w1cog_100  + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book  + home.has.internet | schids, data = d_cl) 
did_adj_cog_both <- feols(cog_dif ~ migr_both + w1cog_100  + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book  + home.has.internet | schids, data = d_cl)
```

Results
```{r}
models_adj_cog <- list(
'Any Parent' = did_adj_cog,
'Only Mother' = did_adj_cog_m_only,
#'Mother' = #did_adj_cog_m,
'Only Father' = did_adj_cog_f_only,
#'Father' = #did_adj_cog_f,
'Both Parents' = did_adj_cog_both) 
```

```{r}
table_adj_cog <- models_adj_cog |> msummary(
  stars =  stars,
  fmt = 2,
  title = title_cog,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  output = 'latex'
  )
#table_adj_cog
```

Regressing on total exam score
```{r}
did_adj_ttl <- feols(avg_dif ~ migr + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = d_cl) 
did_adj_ttl_m_only <- feols(avg_dif ~ migr_mo_only + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = d_cl) 
#did_adj_ttl_m <- feols(avg_dif ~ migr_mo + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = d_cl) 
#did_adj_ttl_f <- feols(avg_dif ~ migr_fa + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = d_cl) 
did_adj_ttl_f_only <- feols(avg_dif ~ migr_fa_only + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = d_cl) 
did_adj_ttl_both <- feols(avg_dif ~ migr_both + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = d_cl)
```

Results
```{r}
models_adj_ttl <- list(
'Any Parent' = did_adj_ttl,
'Only Mother' = did_adj_ttl_m_only,
#'Mother' = #did_adj_ttl_m,
'Only Father' = did_adj_ttl_f_only,
#'Father' = #did_adj_ttl_f,
'Both Parents' = did_adj_ttl_both)
```

```{r}
table_adj_ttl <- models_adj_ttl |> msummary(
  stars =  stars,
  fmt = 2,
  title = title_ttl,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  output = 'latex'
  )
#table_adj_ttl
```

Regressing on total exam score
```{}
did_adj_ttl_z <- feols(ttl_z_dif ~ migr + w1ttl_z + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = d_cl) 
did_adj_ttl_z_m_only <- feols(ttl_z_dif ~ migr_mo_only + w1ttl_z + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = d_cl) 
did_adj_ttl_z_m <- feols(ttl_z_dif ~ migr_mo + w1ttl_z + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = d_cl) 
did_adj_ttl_z_f <- feols(ttl_z_dif ~ migr_fa + w1ttl_z + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = d_cl) 
did_adj_ttl_z_f_only <- feols(ttl_z_dif ~ migr_fa_only + w1ttl_z + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = d_cl) 
did_adj_ttl_z_both <- feols(ttl_z_dif ~ migr_both + w1ttl_z + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = d_cl)
```

Results
```{}
models_adj_ttl_z <- list(
'Any Parent' = did_adj_ttl_z,
'Only Mother' = did_adj_ttl_z_m_only,
#'Mother' = did_adj_ttl_z_m,
'Only Father' = did_adj_ttl_z_f_only,
#'Father' = did_adj_ttl_z_f,
'Both Parents' = did_adj_ttl_z_both) 
table_adj_ttl_z <- models_adj_ttl_z |> msummary(stars = c('*' = .05, '**' = .01, '***' = .001), fmt = 2, coef_rename =  c("migr" = "Migration", "migr_mo_only" = "Migration", "migr_mo" = "Migration", "migr_fa" = "Migration", "migr_fa_only" = "Migration", "migr_both" = "Migration", "w1ttl_dif" = "Baseline difference"), output = 'latex')
table_adj_ttl_z
```

Regressing on total exam score
```{}
did_adj_ttl_dm <- feols(ttl_dm_dif ~ migr + w1avg_100_dm + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = d_cl) 
did_adj_ttl_dm_m_only <- feols(ttl_dm_dif ~ migr_mo_only + w1avg_100_dm + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = d_cl) 
did_adj_ttl_dm_m <- feols(ttl_dm_dif ~ migr_mo + w1avg_100_dm + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = d_cl) 
did_adj_ttl_dm_f <- feols(ttl_dm_dif ~ migr_fa + w1avg_100_dm + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = d_cl) 
did_adj_ttl_dm_f_only <- feols(ttl_dm_dif ~ migr_fa_only + w1avg_100_dm + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = d_cl) 
did_adj_ttl_dm_both <- feols(ttl_dm_dif ~ migr_both + w1avg_100_dm + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = d_cl)
```

Results
```{}
models_adj_ttl_dm <- list(
'Any Parent' = did_adj_ttl_dm,
'Only Mother' = did_adj_ttl_dm_m_only,
#'Mother' = did_adj_ttl_dm_m,
'Only Father' = did_adj_ttl_dm_f_only,
#'Father' = did_adj_ttl_dm_f,
'Both Parents' = did_adj_ttl_dm_both) 
table_adj_ttl_dm <- models_adj_ttl_dm |> msummary(stars = c('*' = .05, '**' = .01, '***' = .001), fmt = 2, coef_rename =  c("migr" = "Migration", "migr_mo_only" = "Migration", "migr_mo" = "Migration", "migr_fa" = "Migration", "migr_fa_only" = "Migration", "migr_both" = "Migration", "w1ttl_dif" = "Baseline difference"), output = 'latex')
table_adj_ttl_dm
```

## Matching
```{r}
m.out <- matchit(
  migr ~ w1cog_100 + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.is.ethnic.minority + child.education.aspiration + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet, 
  data = d_cl, 
  #replace = T,
  ratio = 3,
  method = "optimal", 
  distance = "glm",
  exact = migr ~ ctyids
  )
```

```{r}
sum_m.out <- summary(m.out)
fig_msmd_any <- plot(sum_m.out, xlim = c(0, 0.55), position = 'right')
```

```{r}
dm <- match.data(m.out)
```

```{r}
didm_cog <- feols(cog_dif ~ migr + w1cog_100 | schids, data = dm, weights = dm$weights)

didm_ttl <- feols(avg_dif ~ migr + w1avg_100 | schids, data = dm, weights = dm$weights)

didm_adj_cog <- feols(cog_dif ~ migr + w1cog_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = dm, weights = dm$weights)

didm_adj_ttl <- feols(avg_dif ~ migr + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet  | schids, data = dm, weights = dm$weights)

```

```{r}
d_mo_only <- d_cl |> drop_na(migr_mo_only)
m.out_mo_only <- 
  matchit(migr_mo_only ~ w1cog_100 + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.is.ethnic.minority + child.education.aspiration + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet, 
  data = d_mo_only, 
  #replace = T,
  method = "optimal",
  distance = "glm",
  ratio = 3,
  exact = migr_mo_only ~ ctyids
  )
```

```{r}
sum_m.out_mo_only <- summary(m.out_mo_only)
fig_msmd_mo_only <- plot(sum_m.out_mo_only, xlim = c(0, 0.55), position = 'right')
```

```{r}
dm_mo_only <- match.data(m.out_mo_only)
```

```{r}
didm_cog_mo_only <- feols(cog_dif ~ migr_mo_only + w1cog_100 | schids, data = dm_mo_only, weights = dm_mo_only$weights)

didm_ttl_mo_only <- feols(avg_dif ~ migr_mo_only + w1avg_100 | schids, data = dm_mo_only, weights = dm_mo_only$weights)

didm_adj_cog_mo_only <- feols(cog_dif ~ migr_mo_only + w1cog_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = dm_mo_only, weights = dm_mo_only$weights)

didm_adj_ttl_mo_only <- feols(avg_dif ~ migr_mo_only + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet  | schids, data = dm_mo_only, weights = dm_mo_only$weights)

```

```{r}
d_fa_only <- d_cl |> drop_na(migr_fa_only)
m.out_fa_only <- matchit(
  migr_fa_only ~ w1cog_100 + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.is.ethnic.minority + child.education.aspiration + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet,
  data = d_fa_only,
  #replace = T,
  method = "optimal",
  distance = "glm",
  ratio = 3,
  exact = migr_fa_only ~ ctyids
  )
```

```{r}
sum_m.out_fa_only <- summary(m.out_fa_only)
fig_msmd_fa_only <- plot(sum_m.out_fa_only, xlim = c(0, 0.55), position = 'right')
```

```{r}
love.plot(
  m.out,
  stats = "variance.ratios",
  binary = "std", 
  #limits = c(0, 0.55),
  )
love.plot(
  m.out_mo_only,
  stats = "variance.ratios",
  binary = "std", 
  #limits = c(0, 0.55),
  )
love.plot(
  m.out_fa_only,
  stats = "variance.ratios",
  binary = "std", 
  #limits = c(0, 0.55),
  )
love.plot(
  m.out_both,
  stats = "variance.ratios",
  binary = "std", 
  #limits = c(0, 1.5),
  )
```

```{r}
dm_fa_only <- match.data(m.out_fa_only)
```

```{r}
didm_cog_fa_only <- feols(cog_dif ~ migr_fa_only + w1cog_100 | schids, data = dm_fa_only, weights = dm_fa_only$weights)

didm_ttl_fa_only <- feols(avg_dif ~ migr_fa_only + w1avg_100 | schids, data = dm_fa_only, weights = dm_fa_only$weights)

didm_adj_cog_fa_only <- feols(cog_dif ~ migr_fa_only + w1cog_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = dm_fa_only, weights = dm_fa_only$weights)

didm_adj_ttl_fa_only <- feols(avg_dif ~ migr_fa_only + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet  | schids, data = dm_fa_only, weights = dm_fa_only$weights)

```

```{r}
d_both <- d_cl |> drop_na(migr_both) 
m.out_both <- matchit(
  migr_both ~ w1cog_100 + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.is.ethnic.minority + child.education.aspiration + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level +  home.economic.resource + home.extracurricular.book + home.has.internet, 
  data = d_both, 
  method = "optimal",
  distance = "glm",
  #replace = T,
  ratio = 3,
  exact = migr_both ~ ctyids
  )
```

```{r}
plot(summary(m.out_both), xlim = c(0, 0.55), position = 'right')
```
```{}
love.plot(
  m.out_both,
  threshold = c(m = .1),
  binary = "std", 
  limits = c(0, 0.55),
  abs = T
  )
love.plot(
  m.out_both,
  stats = "variance.ratios",
  binary = "std", 
  #limits = c(0, 0.55),
  abs = T
  )

```

```{r}
dm_both <- match.data(m.out_both)
```

```{r}
didm_cog_both <- feols(cog_dif ~ migr_both + w1cog_100 | schids, data = dm_both, weights = dm_both$weights)

didm_ttl_both <- feols(avg_dif ~ migr_both + w1avg_100 | schids, data = dm_both, weights = dm_both$weights)

didm_adj_cog_both <- feols(cog_dif ~ migr_both + w1cog_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = dm_both, weights = dm_both$weights)

didm_adj_ttl_both <- feols(avg_dif ~ migr_both + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet  | schids, data = dm_both, weights = dm_both$weights)

```

```{}
d_mo <- d_cl |> drop_na(migr_mo)
m.out_mo <- matchit(
  migr_mo ~ w1cog_100 + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet, data = d_mo, 
  method = "optimal",
  #replace = T,
  ratio = 3,
  exact = migr_mo ~ ctyids
  )
summary(m.out_mo, un = FALSE)
```

```{}
dm_mo <- match.data(m.out_mo)
```

```{}
feols(cog_dif ~ migr_mo + w1cog_100| schids, data = dm_mo, weights = dm_mo$weights)

feols(avg_dif ~ migr_mo + w1avg_100  | schids, data = dm_mo, weights = dm_mo$weights)

feols(cog_dif ~ migr_mo + w1cog_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = dm_mo, weights = dm_mo$weights)

feols(avg_dif ~ migr_mo + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet  | schids, data = dm_mo, weights = dm_mo$weights)

```

```{}
d_fa <- d_cl |> drop_na(migr_fa)
m.out_fa <- matchit(
  migr_fa ~ w1cog_100 + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet, data = d_fa,
  method = "optimal",
  #replace = T,
  ratio = 3,
  exact = migr_fa_only ~ ctyids
  )
summary(m.out_fa, un = FALSE)
```

```{}
dm_fa <- match.data(m.out_fa)
```

```{}
feols(cog_dif ~ migr_fa + w1cog_100 | schids, data = dm_fa, weights = dm_fa$weights)

feols(avg_dif ~ migr_fa + w1avg_100 | schids, data = dm_fa, weights = dm_fa$weights)

feols(cog_dif ~ migr_fa + w1cog_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = dm_fa, weights = dm_fa$weights)

feols(avg_dif ~ migr_fa + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet  | schids, data = dm_fa, weights = dm_fa$weights)

```

```{r}
title_cog_matching <- 'Parental migration’s effect on children’s cognitive test scores, estimated with matching'
title_ttl_matching <- 'Parental migration’s effect on children’s academic exam scores, estimated with matching'
note_matching <- "3:1 optimal pair propensity score matching; exact matching by county"
```

```{r}
models_m_cog <- list(
'Any Parent' = didm_cog,
'Only Mother' = didm_cog_mo_only,
##'Mother' = #did_adj_cog_m,
'Only Father' = didm_cog_fa_only,
##'Father' = #did_adj_cog_f,
'Both Parents' = didm_cog_both) 
```

```{r}
table_m_cog <- models_m_cog |> msummary(
  stars =  stars,
  fmt = 2,
  title = title_cog_matching,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  notes = note_matching,
  output = 'latex'
  )
#table_m_cog
```

```{r}
models_m_ttl <- list(
'Any Parent' = didm_ttl,
'Only Mother' = didm_ttl_mo_only,
##'Mother' = #did_adj_ttl_m,
'Only Father' = didm_ttl_fa_only,
##'Father' = #did_adj_ttl_f,
'Both Parents' = didm_ttl_both) 
```

```{r}
table_m_ttl <- models_m_ttl |> msummary(
  stars =  stars,
  fmt = 2,
  title = title_ttl_matching,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  notes = note_matching,
  output = 'latex'
  )
#table_m_ttl
```

```{r}
models_m_adj_cog <- list(
'Any Parent' = didm_adj_cog,
'Only Mother' = didm_adj_cog_mo_only,
##'Mother' = #did_adj_cog_m,
'Only Father' = didm_adj_cog_fa_only,
##'Father' = #did_adj_cog_f,
'Both Parents' = didm_adj_cog_both) 
```

```{r}
table_m_adj_cog <- models_m_adj_cog |> msummary(
  stars =  stars,
  fmt = 2,
  title = title_cog_matching,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  notes = note_matching,
  output = 'latex'
  )
#table_m_adj_cog
```

```{r}
models_m_adj_ttl <- list(
'Any Parent' = didm_adj_ttl,
'Only Mother' = didm_adj_ttl_mo_only,
##'Mother' = #did_adj_ttl_m,
'Only Father' = didm_adj_ttl_fa_only,
##'Father' = #did_adj_ttl_f,
'Both Parents' = didm_adj_ttl_both) 
```

```{r}
table_m_adj_ttl <- models_m_adj_ttl |> msummary(
  stars =  stars,
  fmt = 2,
  title = title_ttl_matching,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  notes = note_matching,
  output = 'latex'
  )
#table_m_adj_ttl
```

```{r}
num.outlier <- d_cl%>%filter(w1mat_100>100|w1eng_100>100|w2mat_100>100|w2eng_100>100)%>%count
```

## Interaction
```{}
d_cl1 <- d_cl |> mutate(
  home.economic.resource.w2 = if_else(is.na(w2be23), w2a09, w2be23)
) |> mutate(
  fin.better = if_else(home.economic.resource.w2 > home.economic.resource,
                       1,
                       0)
  )

```

```{}
feols(avg_dif ~ migr*fin.better + w1avg_100 | schids, data = d_cl1) |>
  msummary(stars = c('†' = .1, '*' = .05, '**' = .01, '***' = .001),
           output = 'latex')
feols(cog_dif ~ migr*fin.better + w1cog_100 | schids, data = d_cl1)  |>
  msummary(stars = c('†' = .1, '*' = .05, '**' = .01, '***' = .001),
           output = 'latex')
feols(avg_dif ~ migr_mo_only*fin.better + w1avg_100 | schids, data = d_cl1) |>
  msummary(stars = c('†' = .1, '*' = .05, '**' = .01, '***' = .001),
           output = 'latex')
feols(cog_dif ~ migr_mo_only*fin.better + w1cog_100 | schids, data = d_cl1)  |>
  msummary(stars = c('†' = .1, '*' = .05, '**' = .01, '***' = .001),
           output = 'latex')
feols(avg_dif ~ migr_fa_only*fin.better + w1avg_100 | schids, data = d_cl1) |>
  msummary(stars = c('†' = .1, '*' = .05, '**' = .01, '***' = .001),
           output = 'latex')
feols(cog_dif ~ migr_fa_only*fin.better + w1cog_100 | schids, data = d_cl1)  |>
  msummary(stars = c('†' = .1, '*' = .05, '**' = .01, '***' = .001),
           output = 'latex')
feols(avg_dif ~ migr_both*fin.better + w1avg_100 | schids, data = d_cl1) |>
  msummary(stars = c('†' = .1, '*' = .05, '**' = .01, '***' = .001),
           output = 'latex')
feols(cog_dif ~ migr_both*fin.better + w1cog_100 | schids, data = d_cl1)  |>
  msummary(stars = c('†' = .1, '*' = .05, '**' = .01, '***' = .001),
           output = 'latex')
```

```{}
feols(avg_dif ~ migr*fin.better + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet  | schids, data = d_cl1) |>
    msummary(stars = c('†' = .1, '*' = .05, '**' = .01, '***' = .001),
           output = 'latex')
feols(cog_dif ~ migr*fin.better + w1cog_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet  | schids, data = d_cl1) |>
    msummary(stars = c('†' = .1, '*' = .05, '**' = .01, '***' = .001),
           output = 'latex')
```

```{}
d_cl1=d_cl1 |> mutate(
  supervision.less = if_else((w2a18 + w2a19) < (b2201 + b2202), 1, 0)
)
```

```{}
feols(avg_dif ~ migr*supervision.less + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet  | schids, data = d_cl1) |>
    msummary(stars = c('†' = .1, '*' = .05, '**' = .01, '***' = .001),
           output = 'latex')
feols(cog_dif ~ migr*supervision.less + w1cog_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + mother.education.level + father.education.level + home.economic.resource + home.extracurricular.book + home.has.internet  | schids, data = d_cl1) |>
    msummary(stars = c('†' = .1, '*' = .05, '**' = .01, '***' = .001),
           output = 'latex')
```

```{r}
bal.tab(m.out)
```