### 2022-07-05 appendix.Rmd

# Tables

```{r tab-reg-cog-adj}
table_adj_cog
```

```{r tab-reg-ttl-adj}
table_adj_ttl
```

```{r tab-reg-m-adj-cog}
table_m_adj_cog
```

```{r tab-reg-m-adj-ttl}
table_m_adj_ttl
```

\newpage