### 2022-07-08 MScThesisText.Rmd

---
title             : "Newly Left-Behind Children's Educational Performance Drops Slightly After Parental Migration"
shorttitle        : "Left-Behind Children"

author: 
  - name          : "Jiancheng Gu"
    affiliation   : "1"
    address       : "De Boelelaan 1105, 1081 HV - Amsterdam, The Netherlands"
    email         : "j4.gu@student.vu.nl"

affiliation:
  - id            : "1"
    institution   : "Faculty of Social Sciences, Vrije Universiteit Amsterdam"

authornote: |

abstract: |
  Millions of children worldwide are "left behind" at home communities as their parents migrate away for work. 
  Parents' labor migration increases household income but decreases parental care, thereby exerting mixed influences on child development outcomes like educational attainment. 
  To what extent do different arrangements of parental migration affect children's educational performance in the short term? 
  Studying the case of China, I obtained a sample of four thousand junior high school students from a nationally representative two-wave panel survey. 
  I differentiated household migration arrangements of any one parent, of mother only, of father only, and of both parents.
  I measured children's schooling performance by cognitive test scores and academic exam grades.
  For causal inference, I built difference-in-differences models with propensity score matching. 
  Results show that the newly left-behind children's performance at school dropped slightly compared to that of non-left-behind peers. 
  Mother-only migration is negatively associated with children’s academic grades and cognitive abilities. 
  Father-only migration is negatively associated with children’s cognitive abilities, but its magnitude is greater than mother-only migration.

keywords          : "labor migration, left-behind children, educational performance"
wordcount         : "6,600"

floatsintext      : yes
linenumbers       : no
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

header-includes:
  - |
    \makeatletter
    \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-1em}%
      {\normalfont\normalsize\bfseries\typesectitle}}

    \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-\z@\relax}%
      {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
    \makeatother

    \usepackage{float}
    \usepackage{booktabs}
    \usepackage{siunitx}
    \usepackage{setspace}
    \onehalfspacing

csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass     : "apa7"

classoption       : "man"
output            : papaja::apa6_pdf
appendix: 
  - "appendix.Rmd"
  - "appendix-b.Rmd"
bibliography: references.bib
---

```{r setup, include = FALSE}
load('.Rdata')
library("papaja")
library("cobalt")
library("kableExtra")
library("modelsummary")
options(modelsummary_get="broom") 
r_refs("references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Introduction

Globally, millions of children reside in their home communities but live apart from their parents as either one or both parents have migrated for work. These are the so-called "left-behind children" or "stay-behind children." They are common in migrant-sending regions, such as Latin America, Sub-Sahara Africa, East Europe, and large parts of Asia. Researchers estimated that one-sixth of Mexican children [@dewaard2018], one-fifth of Bulgarian children [@popova2018], and one-fourth of Chinese children were separated from their migrating parents [@nbs2017]. The issue of left-behind children has caught the attention of policymakers and legislators. In a resolution, the Council of Europe's @parliamentaryassembly2021 expressed concern about "the scale of this phenomenon and the long-term damage it creates" and declared that leaving "millions of labour migrants' children without parental care is a mass violation of human rights".

How are stay-behind children faring in the absence of parents? Research findings are mixed due to the differential effects of parental migration. Consider children's education. According to the review articles by @debrauw2019 and @vanhook2020, labor migrants send remittances back, which can fund children's schooling as well as improve their nutritional and living conditions. To what degree can the inflow of remittances compensate for the loss of parental attention? Parental migration decreases the amount and quality of parental guidance, protection, and support. It is uncertain how these two opposing effects together shape children's educational performance. This uncertainty merits empirical investigation.

A challenge arises in estimating the effect of parent's migration on children's educational achievement. Experiments can unveil causal relations. It is unfeasible and unethical to design an experiment that randomly assigns parents to migrate without their children. Alternatively, a quasi-experimental design can help with causal inference. It requires that the cause be on par with random assignment to participants, conditional on identification assumptions [@cunningham2021].

My research question is: To what extent do different arrangements of parental migration affect the educational performance of left-behind children in the short term? I choose China as the case and analyze data from a two-wave panel survey. The nationally representative sample includes children in both urban and rural areas. I develop a difference-in-differences model for causal inference and apply propensity score matching for robustness checks. The outcome is measured with children's cognitive test score and academic exam grade. I demonstrate that parental migration reduces the educational performance of children staying behind, but only to a small extent.

## China's Context

Three features make China a suitable case for studying left-behind children. The first feature is the huge scale of internal labor migration. According to the National Bureau of Statistics, the year 2021 recorded 292.5 million in-country migrant workers. They form the backbone of essential industries, such as retail, transportation, and manufacturing. The 2015 one-percent National Population Sample Survey estimated that 68.7 million Chinese children were left behind [@nbs2022; @nbs2017].

The internal migration stemmed from the neoliberal economic regime. Scholars like @duckett2020 and @zhan2020 argued that China conforms to the neoliberal trend, extending the logic of the market to other domains of human life. In the 1980s, the ruling communist party ditched the planned economy, initiated market reforms, and opened up the country. Since then, China has integrated in the global economy and contributed substantial growth to it. The economic reform and globalization have enriched the coastal cities like Shanghai and Shenzhen. The rural hinterlands, however, have lagged behind. Simultaneously, many people have left the countryside for the city. The regional inequality needs social policy to amend. But China's welfare program, as @qian2021 argued, is fragmented and stratified. People who reap the greatest benefits are public servants, urban dwellers, and formally employed workers. As rural migrants arrive in the city, they find themselves with limited access to social welfare, such as health care and children's education [@qian2021]. The welfare system's exclusion of rural migrants means that the state shifts its responsibility to individuals, resembling the neoliberal governmentality [@zhan2020].

The institutional discrimination against labor migrants finds its roots in China's "multilevel citizenship" system. As @vortherms2021 explained, the constituents of citizenship -- "membership rules and rights entitlements" -- operate below the national level in China. The local state confers most of the rights and defines who is eligible to these rights through *hukou*, or the household registration system:

> A political institution rooted in centuries of Chinese bureaucracy, the modern hukou is the primary identity document defining membership and entitlements for local citizenship. Accessing fundamental citizenship rights depends on your registration status ... Registration includes basic identifying information, a specific address of registration, and a type: urban, rural, or resident. [@vortherms2021]

In other words, the hukou system binds one's citizenship rights with one's place of registry. In theory, Chinese citizens can migrate freely within the country. In practice, they cannot change hukou freely. When rural residents move to the city, it takes months -- if not years -- to apply for a new hukou. Before receiving urban hukou, they can hardly send their children into public schools in the host city [@gu2022]. Migrant workers have organized petitions, strikes, and protests. These actions often fell on deaf ears and sometimes even ended with police crackdown [@chan2019]. Facing financial and policy constraints, many migrant laborers have to leave children in their hometowns.

How does China counter the issue of labor migration and left-behind children? According to @wang2020 and @zhang2018, the government has gradually liberalized the multilevel citizenship scheme. It has introduced residence permit for internal migrants as the basis for social entitlements. First-tier cities like Beijing and Shanghai have brought forth a point-based system to grant hukou to newcomers. The system's evaluation criteria include in-demand talents and skills, educational and professional qualifications, and consecutive years of employment and tax-payment. This parallels how European states differentially incorporate foreign immigrants into national welfare regimes [@zhang2018]. Another policy response, as stated by @roberts2018, is the centralized boarding school. The Ministry of Education promotes boarding schools in the rural county-seats. These schools serve as a cost-effective way to enroll students in surrounding villages and to look after the left-behind children. In reality, such schools "are often ill-equipped to ensure the healthy upbringing and psychological wellbeing of children" (p. 6).

Social scientists have researched extensively about China's left-behind children, but the current literature still has three major shortcomings. The first is that most literature focuses on children left behind in rural areas while neglecting the urban counterparts. The National Bureau of Statistics stated that the number of urban left-behind children had been growing and expected that the trend would continue [@nbs2017]. The second shortcoming lies in sampling and research design. Some studies suffer from a limited sample size of less than one thousand participants or a limited geographic coverage of one to a few towns [@liu2021; @yang2022; @shu2021]. Others only estimated correlation using cross-sectional data [@akezhuoli2022; @jin2020; @huang2020]. These studies could have employed causal inference techniques like fixed effect, matching, or inverse probability of treatment weighting. The third shortcoming lies in the measurement of outcome. @chang2019 and @wang2019 measured students' educational performance with the standardized test score of only one academic subject, namely mathematics. To tackle these limitations, I adopted a nationally representative two-wave panel dataset. My sample includes urban left-behind children as well. I included the test results of multiple subjects to measure comprehensive academic achievement.

## Theory and Hypotheses

Between parental migration and children's education, the mechanisms trace to two theoretical frameworks. One is the resource generation model. It holds that parental migration brings forth greater resources in the form of economic capital. Economic capital is, in Bourdieu's words, "immediately and directly convertible into money and may be institutionalized in the form of property rights" [-@bourdieu2002, p.16]. Economic capital is transferable to human capital, which is the stock of "skills, talents, health, and expertise" in people that enhance their productivity [@goldin2016; @botev2019]. @yang2008 noted that in the Philippines, a sharp depreciation of peso increased the amount of remittances from migrants working abroad. This subsequently boosted children's school enrollment and educational expenditure. @cebotari2020 developed a fixed effect model with longitudinal data collected in Ghana. He found that stay-behind children's education benefited from "the steady inflow of both in-kind and monetary remittances" contributed by migrant parents. In China, @chang2019 interviewed migrant workers and their family members. Oftentimes, labor migration presented the only way to sustain their children's education and livelihood. @peng2018 interviewed first-generation migrant mothers whose initial migration dated back to the 1990s. To them, "good mothering" primarily meant offering economic support for their children in education, which can help their children climb the ladder of success.

The other theoretical model is the family disruption model. Children flourish under parental nurturing care. This term refers to the "stable environments that promote health and adequate nutrition, protect from threats ... opportunities for learning," and "responsive, emotionally supportive and developmentally enriching relationships" [@black2021, p. 1]. Parental migration decreases nurturing care, as characterized by less homework supervision and emotional support by parents [@wen2019]. Hong talked to a group of left-behind children at a rural school [@hong2019]. These students expressed the utter sense of loneliness "with no advocate, no support and with no one looking out for them" (p. 13). Many of them felt indifferent to grades and disengaged with school. Some even longed to quit school to pursue a free, independent life (p. 14). Similarly, @murphy2022 talked to left-behind adolescents at a village. Some of them "embraced the 'left-behind' label" because it captured their need for more parental care (p. 185), "especially companionship and affection in daily life" (p. 186).

Given the two opposing effects, researchers have obtained mixed results on how parental migration impacts children's education. @botezat2020 adopted past immigration rate as an instrumental variable. They found a positive effect of parents' migration on academic grades of adolescents aged eleven to fifteen in Romania. The team of @wang2019 collected data from junior high school students in central China. Their two-way fixed-effect model showed no significant effect of parental migration on left-behind children's achievement in mathematical test. @nguyen2016 conducted fixed-effect regression and demonstrated that parental migration was associated with lower cognitive capabilities among Indian and Vietnamese children aged five to eight.

Parental migration immediately disrupts family structure, which in turn lowers academic performance of children. It takes longer time for labor migration to generate resources that improve children's educational attainment. I propose the first hypothesis:

$H_1$ In the short term, parental migration reduces the educational performance of children left-behind.

$H_1a$ In the short term, parental migration negatively affects left-behind children's cognitive test score.

$H_1b$ In the short term, parental migration negatively affects left-behind children's academic exam score.

Researchers have studied whether the impact differ between mother's migration and father's migration. @xu2019 adopted a fixed-effects propensity score weighting model on cross-sectional data collected from seventh and ninth graders in rural areas across China. They found that the absence of father only or both parents had "little or no association with negative outcomes" on children's academic and cognitive performances. Only-mother absence, however, showed a "strong association with negative outcomes" (p. 1646). @chen2019 conducted structural equation modeling on the cross-sectional data collected from fourth to seventh graders in the countryside of one province. They found that maternal migration was "negatively associated with children's social competence and academic performance," while paternal migration had no direct effects on the two outcomes (pp. 860-861).

Scholars like @murphy2020 noted that in some Asian and Latin American societies, the social norms designate men as breadwinners and women as caregivers. Under such norms, fathers' migration is more likely to win children's acceptance than does mothers' migration. Left-behind children may even resent their migrant mothers for transgressing on the expected care-giving roles [@murphy2020]. I propose the second hypothesis:

$H_2$ In the short term, mother's migration more negatively affect the educational performance of children left-behind than do father's migration.

$H_2a$ In the short term, mother's migration more negatively affect left-behind children's cognitive test score than do father's migration.

$H_2b$ In the short term, mother's migration more negatively affect left-behind children's academic exam score than do father's migration.

\newpage

# Methods

## Sampling

I obtained the data from the China Education Panel Survey (CEPS) [@nsrc]. It is a nationally representative, school-based survey featuring junior high school students. Currently, two waves of data are available. The CEPS team operates under the National Survey Research Center at the Renmin University of China. The survey team adopted multi-stage stratified probability proportional to size sampling and administered a paper-and-pencil questionnaire to each participant. The wave one baseline survey took place in the 2013-2014 academic year with a sample size of about twenty thousand students in the seventh and ninth grades. They were nested in 438 classrooms of 112 schools in 28 urban districts or rural counties (hereafter 'counties'). The wave two follow-up survey in the 2014-2015 academic year lost track of the ninth grade cohort and only surveyed `r printnum(nrow(d))` of the seventh grade cohort. I further restricted the sample to students living with both parents at the original domicile (place of hukou registry) at the baseline survey. To tackle missing values, I adopted list-wise deletion for the dependent, independent, and control variables. The final sample size amounts to `r printnum(nrow(d_cl))`.

## Measurement

I explored two categories of educational performance: cognitive and academic abilities. One key dependent variable, the student's cognitive skill, was measured by a fifteen-minute standardized test administered by the CEPS research team. The test covered three types of reasoning: verbal, visuospatial, and numerical. The baseline and follow-up tests consisted of 20 and 35 items, respectively. The raw score was calculated based on the Item Response Theory [@wang2015]. I transformed the raw scores to the 100-point scale. I chose the raw score instead of the z-score, as z-standardization fails to document the change of mean between time points, which is also of interest [@moeller2015].

Another dependent variable concerned students' academic performance. This was measured by the average scores of mid-term exams in Chinese, mathematics, and English as provided by each school. Exam scores are reliable indicators of learning performance, and are comparable across schools for the same cohort of students. This is related to how junior high schools operate within the compulsory education system in China. According to @oecd2016, China's junior secondary education is a unified system that does not distinguish between academic and vocational tracks. At school, the subjects of Chinese, mathematics, and English are core subjects and others as supplementary subjects. Throughout junior high school, students study the core subjects in preparation for the senior high school entrance exam. Within each cohort in one particular school, teachers use the same syllabus and administer the same exams during a given assessment period. Across cities and regions, junior high schools adopt the educational content and assessment set by the Ministry of Education [@oecd2016].

I transformed the exam grades through the following steps. I first aligned different grading scales (full mark of 100, 120, and 150 for one subject) with the 100-point scale. After the conversion, I checked for abnormal values lower than zero or higher than one hundred. `r printnum(nrow(outlier))` observations were outliers, accounting for less than one percent of the sample. I retained these observations in the sample and trimmed the subject exam scores. The trimming ensured that all values would be within the 0-100 range. In this way, I can limit the impact of outliers. Finally, I calculated the average score of these three subjects.

The key independent variable, the parental migration status, was identified by two items in the parent survey. In both the baseline and the follow-up surveys, parents specified the family members living in the same household at the time. I constructed four treatment dummy variables measuring migration arrangements: any parents migrated, only father migrated, only mother migrated, and both parents migrated. The latter three arrangements are mutually exclusive. The control group consisted of parents staying with their children throughout the two waves of survey. I then removed the observations that reported parental divorce or death. These type of parental absence are irrelevant to migration.

The control variables at children's level included age, gender, self-rated health, hukou type, number of siblings, whether the child is ethnic minority of Han Chinese, whether the child had skipped grade or repeated grade, and the level of education that the child aspires to achieve. At the household level, I controlled the levels of education of father and mother, the parental expectation on the child's level of education, whether the parent has a white collar job, whether the household has internet access, the perceived household financial conditions, and the number of extracurricular books at home. The lists of covariates are in Table \@ref(tab:tab-ctvar-dit) -- Table \@ref(tab:tab-ctvar-msr) in the Results section.

## Analytical Strategy

I built a two-way fixed effect model in a difference-in-differences (DID) design, following Bai and colleagues' [-@bai2018; -@bai2020] approach. The model served to analyze the educational outcome of children newly left behind by parents vis-a-vis that of children staying together with parents. I first specified an unrestricted and unadjusted model: $$\Delta score_{i,j} = \alpha + \beta \cdot migr_{i} + \gamma \cdot score_{i;1} + \lambda \cdot S_{j} + \varepsilon_{i,j}$$ For student $i$ in school $j$, $\Delta score_{i,j}$ is the change in test score between the two waves of surveys, $migr_{i}$ is the treatment dummy variable, $score_{i,j;1}$ is the test score at baseline, and $S_{j}$ is the school fixed-effect. School fixed-effect can account for differences due to time-invariant factors at the school and county levels, as students nested in one school are also in the same county. An alternative is to use a class fixed-effect to control for time-invariant factors at the classroom level and above. However, about one-sixth of the students were reassigned into new classes at the end of baseline academic year. As the reassignment of classroom will complicate the model, I ruled out the classroom fixed-effect.

The DID model is unrestricted because it does not restrict on the coefficient associated with the baseline scores. The model is unadjusted as it does not adjust for additional covariates. Theoretically, it is unnecessary to include covariates that vary over group but remain constant over time; They would cancel out in the two-way fixed effect model [@huntington-klein2022]. The standard errors are clustered at the classroom level.

In addition, I made an unrestricted and adjusted DID model: $$\Delta score_{i,j} = \alpha + \beta \cdot migr_{i} + \gamma \cdot score_{i;1} + \lambda \cdot S_{j} + \theta \cdot X_{i} + \varepsilon_{i,j}$$ where $X_{i}$ is the vector of covariates capturing the characteristics of children, their parents, and their households. The control variables were measured in the baseline survey. This version of model lifts the restriction that covariates from the baseline survey would be associated with a coefficient that equals one.

I adopted `R` (version 4.2.0) for all the analyses [@R-base]. This involved the following packages: `tidyverse` (version 1.3.1) for data cleaning [@R-tidyverse]; `fixest` (version 0.10.4) for econometric modeling [@R-fixest]; `MatchIt` (version 4.4.0) and `cobalt` (version 4.3.2) for matching [@R-MatchIt; @R-cobalt]; `modelsummary` (version 1.0.0), `vtable` (version 1.3.3), and `kableExtra` (version 1.3.4) for presenting the results [@R-modelsummary; @R-vtable; @R-kableExtra]; and `papaja` (version 0.1.0) for typesetting [@R-papaja].

\newpage

# Results

## Descriptive Statistics

Table \@ref(tab:tab-idpvar) summarizes the independent variables, namely the migration status reported in the wave two survey. All `r printnum(nrow(d_cl))` children were living with parents in one household in baseline, but about sixteen percent saw at least one parent migrated between the two waves of survey. I call these students "newly left-behind children" and their households "new migration households." In these new migration households, only-mother migration and only-father migration each accounted for two out of five cases. The remaining one-fifth were both-parents migration.

\renewcommand{\arraystretch}{0.72}

```{r tab-idpvar}
datasummary_df(
  data = tab_idpvar,
  output = 'markdown',
  title = 'Independent variables by parental migration'
)
```

\newpage

The dependent variables are summarized by the control group versus the treatment group in Table 2 below and Figure \@ref(fig:v-plot-cog) -- Figure \@ref(fig:v-plot-ttl-1) in the appendix. Compared to non-left-behind peers, newly left-behind children scored lower in wave two's cognitive and academic tests; Their cognitive test score increased less and their academic exam score dropped more. T-tests show that for the baseline outcomes, the differences of the mean values between-group are probably due to chance. For the wave two outcomes and the between-wave changes, the differences of the mean value are statistically significant.

```{r tab-dpvar}
tab_dpvar |> 
  kbl(booktabs = T, caption = 'Dependent variables by parental migration', row.names = F) |>
  add_indent(c(2,3,5,6)) |>
  add_header_above(c(" ", "No = 3422" = 2, "Yes = 640" = 2)) |>
  add_header_above(c(" ", "Parental Migration" = 4)) |>
  footnote(general = '* p < 0.05, ** p < 0.01, *** p < 0.001')
```

\newpage

Table \@ref(tab:tab-ctvar-dit) summarizes the bivariate analysis on the binary control variables. New migration households had lower internet access rate. Among the newly left-behind children, male sex, past grade retention, and boarding at school were more common compared to the non-left-behind peers. Chi-square tests show that these differences in variance are unlikely due to chance.

```{r tab-ctvar-dit}
tab_ctvar_dit |>
  kbl(booktabs = T, caption = 'Dichotomous control variables by parental migration', row.names = F) |>
  add_header_above(c(" ", "No = 3422" = 2, "Yes = 640" = 2)) |>
  add_header_above(c(" ", "Parental Migration" = 4)) |>
  footnote(general = '0 = No, 1 = Yes; * p < 0.05, ** p < 0.01, *** p < 0.001')
```

\newpage

Table \@ref(tab:tab-ctvar-cnt) displays the bivariate analysis for categorical control variables, and Table \@ref(tab:tab-ctvar-msr) documents their measurement scales. While the majority of mothers and fathers had only finished high school or below, they held big expectations for children. Eight out of ten parents wished that their children complete university education. Scholars like @willis2019 explained this pattern with the Chinese social norms and cultural traditions. In ancient China, the Confucian teaching emphasized the cultivation of personhood through learning, diligence, and persistence. In real life, knowledge could translate into power and privilege. Ordinary people could study for the highly competitive *keju* (civil service examination). As the world's "earliest meritocratic institution," this national exam system recruited top performers into the imperial government and the elite class [@chen2020, p. 2030]. Nowadays in China, the emperor and *keju* are long gone, but the influences are long-lasting. Chinese parents hope that their children receive the best education to secure a decent career [@luo2013; @ng2020].

```{r tab-ctvar-cnt}
tab_ctvar_cnt |>
  kbl(booktabs = T, caption = 'Categorical and continuous control variables by parental migration', row.names = F) |>
  add_header_above(c(" ", "No = 3422" = 2, "Yes = 640" = 2)) |>
  add_header_above(c(" ", "Parental Migration" = 4)) |>
  footnote(general = '* p < 0.05, ** p < 0.01, *** p < 0.001')
```

\newpage

```{r tab-ctvar-msr}
datasummary_df(
  data = tab_var,
  title = 'Measurement and distribution of categorical variables',
  output = 'markdown'
)
```

\newpage

## Regression Analysis

I first present the results on children's cognitive test score. The unadjusted DID models show that newly left-behind adolescents are more likely to score lower in cognitive tests than do their non-left-behind counterparts (see Table 6 below). This lends support to Hypothesis 1a. Parental migration is linked to a `r coef(did_cog)[1] |> abs() |> round(2)` points decrease in cognitive test score. Hypothesis 2a is only partially supported. Compared with the non-left-behind peers, children with only father absent are associated with greater decrease (`r coef(did_cog_f_only)[1] |> abs() |> round(2)` points) in cognitive test score than those with only mother absent (`r coef(did_cog_m_only)[1] |> abs() |> round(2)` points). The adjusted DID models produce results similar to the unadjusted models, with one exception (see Table A1 in appendix). That is, for the cognitive score, the both-parent-migrated coefficient's magnitude is lower than that of father-migrated.

```{r tab-reg-cog}
table_cog
```

Next, I show the results on children's academic test grades. In line with Hypothesis 1b, the unadjusted DID models show that newly left-behind adolescents tend to perform worse in school exams than their non-left-behind peers (see Table 7 below). Parental migration is associated with a drop of `r coef(did_ttl)[1] |> abs() |> round(2)` points in academic exam score. Hypothesis 2b finds partial support. The exam grade of only-mother-migrated children sees a `r coef(did_ttl_m_only)[1] |> abs() |> round(2)` points decrease relative to non-left-behind counterparts. In the scenarios of only father migrated and both parents migrated, the results are not statistically significant. The adjusted DID models' results largely resemble those from the unadjusted models (See Table A2 in appendix). Adding control variables measured in the baseline survey does not change the coefficients for parental migration.

\newpage

```{r tab-reg-ttl}
table_ttl
```

Some predictors are strong in both models of cognitive and academic abilities. These are children's age, preschool attendance, and educational aspiration as well as parents' expectation on children. For students, one year older in age is associated with decreases of `r coef(did_adj_cog)[3] |> abs() |> round(2)` points for cognitive test and of `r coef(did_adj_ttl)[3] |> abs() |> round(2)` points for academic exam. Preschool attendance is linked to increases of `r round(coef(did_adj_cog)[8],2)` points for cognitive test and of `r round(coef(did_adj_ttl)[8],2)` points for academic exam. Higher educational expectation from parents and greater educational aspiration from children are both linked to higher scores. The coefficients are `r round(coef(did_adj_cog)[14],2)` points and `r round(coef(did_adj_cog)[12],2)` points for cognitive test, respectively; and `r round(coef(did_adj_ttl)[14],2)` points and `r round(coef(did_adj_ttl)[12],2)` points for academic exams, respectively.

Several predictors are strong in predicting one outcome but not the other. Covariates that only relate to academic grades are children's gender and self-reported health, and the number of extracurricular books at home. Girls outperform boys by `r round(coef(did_adj_ttl)[4],2)` points in academic exams, but do not show advantage in cognitive tests. Students who consider themselves healthier are associated with higher scores in academic exams. On the contrary, those who report having more extracurricular books at home are associated with lower scores in academic exams. Covariates that only relate to cognitive scores is the number of siblings. Having an additional sibling is associated with lower cognitive score of `r coef(did_adj_cog)[7] |> abs() |> round(2)` points.

\newpage

## Matching

Parental migration is not a randomly assigned treatment. A possible circumstance is that smart and ambitious students perform better at school, and high performance and ambition prompt parents to migrate to support children's schooling. Another scenario is that poverty may hinder children's educational achievement while driving parents to migrate. To tackle the selection bias herein, an applicable technique is matching. It aims to two groups comparable by pairing a unit in the treatment group with one or several units in the control group based on observable covariates [@greifer2022].

Popular choices for matching include propensity score matching (PSM) and coarsened exact matching (CEM). As @pan2018 introduced, PSM will first calculate the propensity score, namely the probability of receiving treatment based on the specified covariates. This score serves as a measure of the distance between two units. When handling this distance, PSM can match the units with two different approaches. The nearest-neighbor approach examines all the treated units and selects the control unit closest in distance for pairing. This approach does not try to optimize any condition; each pair of units ignores how other pairing will occur or have occurred. The optimal-pair approach works in a similar fashion but tries to minimize the absolute pairwise distances in total [@pan2018]. Regardless of specific approaches, @king2019 argued against PSM in general as it risks more "imbalance, model dependence, and bias" for already balanced data (p. 444). They recommended CEM as a viable alternative "for data with continuous, discrete, and mixed variables" (p. 450). CEM is superior in achieving balance as it optimizes the entire joint distribution of covariates. The downside is that it leaves many units without a match and discards these unmatched units [@costalli2021].

I performed matching with all the covariates of the adjusted DID model on the four treatments respectively. My initial attempt with CEM left most units unmatched and resulted in a tiny sample. I then made do with PSM. I combined exact matching (based on the student's county of residence) and optimal-pair PSM (based on the other covariates). Exact matching based on county of residence can preserve the original geographic distribution [@bai2018]. The propensity score was estimated using a logistic regression of the treatment on the covariates. The matching ratio of 3:1 means that one treatment unit would pair with up to three control units. According to the matching summary, most of the covariates' absolute standardized mean differences decrease and fall below 0.1 after matching, suggesting an improved overall balance. In appendix, Figure \@ref(fig:cobal-any) -- Figure \@ref(fig:cobal-both) display the absolute standardized mean differences before and after matching.

The results from models with matching strongly resemble the models without matching (see Tables 8 -- 9 below, and Tables A3 -- A4 in appendix). In the models of cognitive test score, I obtained negative, statistically significant coefficients on all treatment variables except that of both-parents-migrated households. In the models of academic exam score, the treatment variables' coefficients are negative and statistically significant for the any-parent-migrated scenario and the only-mother-migrated scenario. The magnitudes of the statistically significant coefficients are similar to those in the models without matching.

```{r tab-reg-m-cog}
models_m_cog |> msummary(
  stars =  stars,
  fmt = 2,
  title = title_cog_matching,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  notes = note_matching,
  output = 'markdown'
  )
```

```{r tab-reg-m-ttl}
models_m_ttl |> msummary(
  stars =  stars,
  fmt = 2,
  title = title_cog_matching,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  notes = note_matching,
  output = 'markdown'
  )
```

\newpage

# Discussion

This thesis infers the causal relation between parents' migration and their stay-behind children's educational performance in the short run. I build difference-in-differences models with fixed-effect to analyze data from the 2013-2014 and 2014-2015 rounds of China Education Panel Survey. Results suggest that parental migration is, in the short run, negatively associated with children's educational performance. This is evidenced by a 1.2 points drop in academic exam grades and a 2.5 points drop in cognitive test scores. This result contrasts with that of DID models by @bai2018: parental migration is linked to higher English test scores for stay-behind children. The discrepancy could be attributed their study's location, a Northwestern province with grinding poverty. There, the positive impact of resource generation probably outweighs the negative impact of family disruption, resulting in an overall positive effect (p. 1165). In my nationwide sample the reverse may hold. The households may be better off, so labor migration's additional income fails to compensate for the care deficit. Another plausibility is that family disruption precedes resource generation. The former's negative effect befalls the left-behind child soon after parental migration, but the latter's positive effect occurs a while later [@gao2019]. Consequently, I obtain a short-term negative effect.

I further examine whether the effect varies across different arrangements of parental migration. Results indicate that in the short term, only father's absence is not linked to children's academic performance; Only mother's absence is negatively associated with children's academic abilities, with a 2.1 points drop in the test score. This is in line with the findings of previous research. For example, @murphy2020 observed that when a Chinese mother migrates, she defies the cultural expectations on maternal care, thereby harboring bitterness in the child's mind. This bitter sentiment can arise even if there is care by father or other relatives. Besides, only-mother migration sometimes indicates deeper issues within the household. The father may be unfit for work or engaged in domestic violence, prompting the mother's labor migration [@murphy2020]. These may all hamper children's educational performance.

Compared to only mother's migration, only father's migration carries a stronger influence on children's cognitive capability, as shown by a further decrease of 1 point in cognitive test score. This echoes the findings of @shen2021 -- in rural China, the father is often a household's best educated adult. His absence means a loss of household human capital, which slows the cognitive development of children. The stay-at-home mother may devote more to the caregiving role. This can partially buffer the impact by her husband's absence and help her child maintain academic performance.

My research has limitations in sampling due to data constraint. Regarding identification, I could only identify parental migration through such a condition: a parent's absence from a household without parental divorce or death. This is because the CEPS questionnaire asked if parents were living in the household, but did not ask why parents were absent. If some parents were jailed or hospitalized, their absence would be wrongly attributed to migration. Besides, I could not control for the migration history of parents, which are not included in the CEPS data. According to @liang2021, some parents might have out-migrated at an earlier date and returned before the baseline survey. Their children would be attributed to the non-left-behind group. However, these children "suffer similar disadvantage as that of left behind children" (p. 12). @liang2021 argued for a new category: the formerly left-behind children.

Limitations also arise from unaccounted endogeneity and exogeneity. As @chen2009 discussed in their paper on left-behind children, endogenous factors may create selection bias. When migration opportunities arise, some parents may worry that leaving children behind would hamper children's education, thereby giving up the opportunity. Some other parents may believe that their children's schooling would not suffer from parental absence, thereby embarking on migration. Parental beliefs of this sort are unavailable in CEPS data. Exogenous factors may create omitted variable bias. Suppose that a natural hazard hit many counties of a province in the sample. This shock can increase parents' out-migration and reduce stay-behind students' grades [@chen2009]. I could not account for such shocks as CEPS data does not include detailed geographic information.

From my research, we can find new directions for scholarly inquiry. Researchers can explore interaction and moderation in the pathways of resource generation and family disruption. When the CEPS team releases the next round of data, researchers can capitalize a the three-wave panel dataset to build multi-period DID models. This would help examine the parallel trend assumption. Novel methods also offer new possibilities. Machine learning, for example, can estimate the heterogeneous effects across subgroups [@liu2021a]. Looking beyond outbound migration, return migration of parents also worth studying [@liang2021]. Not many publications have explored return migrants and their formerly left-behind children in China [@démurger2015; @liu2018].

Let us consider the policy implications. China's multilevel citizenship scheme shall not remain as it is, despite the small reduction of newly left-behind children's schooling performance. The slight drop is only the short-term consequence; The lifelong cumulative effect on children may be extensive [@liang2020; @meng2017]. Besides, educational attainment is only part of human development outcomes. Other outcomes like mental health and non-cognitive skills are also subjected to parental absence's influence [@antia2020]. Policymakers should accelerate the reforms on hukou, opening more doors for parents to migrate together with their children. Besides top-down policymaking, bottom-up approaches are also necessary. School administrators could dedicate more advisory and counseling programs. Social workers could provide paraprofessional training in communities [@wang2020a].

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::

\newpage

# (APPENDIX) Appendix {.unnumbered}

```{r child = "appendix.Rmd"}
```

```{r child = "appendix-b.Rmd"}
```