### 2023-06-09 analysis/Gu-thesis-script.Rmd
Update code, writing, bibliography

---
title: 'parental migration and adolescents educational performance'
output: html_notebook
---

```{r Load packages}
library(tidyverse) 
library(haven)
library(fixest)
library(modelsummary)
library(MatchIt)
library(vtable)
library(labelled)
library(cobalt) 
library(DescTools)
```

## Data cleaning
```{r Read data}
d_stu1 <- read_dta('cepsw1studentEN.dta') #wave 1 student
d_stu2 <- read_dta('cepsw2studentEN.dta') #wave 2 student
d_par1 <- read_dta('cepsw1parentEN.dta') #wave 1 parent
d_par2 <- read_dta('cepsw2parentEN.dta') #wave 2 parent
```

```{r}
d_stu1 <- d_stu1 |> 
  subset(grade9==0) |> 
  select(!c(frame, subsample, sweight, grade9))
d_stu2 <- d_stu2 |> 
  select(!c(w2frame, w2subsample, w2sweight, w1w2sweight))
d_par1 <- d_par1 |> 
  subset(grade9==0) |> 
  select(!c(frame, subsample, grade9)) #Remove wave 1's observations of grade 9; they were not followed up in wave 2
```

```{r Merge data}
d <- d_stu1 |> 
  left_join(d_par1, by=c('ids', 'schids', 'clsids', 'ctyids', 'fall')) |>
  left_join(d_stu2, by=c('ids', 'schids', 'clsids', 'ctyids')) |>
  left_join(d_par2, by=c('ids', 'clsids', 'w2clsids', 'schids', 'ctyids', 'w2fall')) 
rm('d_stu1', 'd_stu2', 'd_par1', 'd_par2')
```

```{r Sample}
d_cl <- d |> 
  filter(w2a0801==1) |> #keep parents who are married (and alive)
  filter(be1201==1 & be1202==1) |> #keep households with both parents staying at baseline
  filter(stmigrant==1) #keep local children 
```

## Variables 
```{r Exam grade conversion}
d_cl <- d_cl |> mutate(
  w1chn_100 = tr_chn/w2upchn*100, #raw exam score at 100-point scale
  w1mat_100 = tr_mat/w2upmat*100,
  w1eng_100 = tr_eng/w2upeng*100,
  w2chn_100 = w2chn/w2upchn*100,
  w2mat_100 = w2mat/w2upmat*100,
  w2eng_100 = w2eng/w2upeng*100
)
```

```{r Outlier detection}
outlier <- d_cl |> filter(
  w1chn_100<0 | w1chn_100>100 | w1mat_100<0 | w1mat_100>100 | w1eng_100<0 | w1eng_100>100 |
  w2chn_100<0 | w2chn_100>100 | w2mat_100<0 | w2mat_100>100 | w2eng_100<0 | w2eng_100>100)
```

```{r Trim outlying values}
d_cl <- d_cl |> mutate(
  w1chn_100_trim = Winsorize(w1chn_100, minval = 0, maxval = 100),
  w2chn_100_trim = Winsorize(w2chn_100, minval = 0, maxval = 100),
  w1mat_100_trim = Winsorize(w1mat_100, minval = 0, maxval = 100),
  w2mat_100_trim = Winsorize(w2mat_100, minval = 0, maxval = 100),
  w1eng_100_trim = Winsorize(w1eng_100, minval = 0, maxval = 100),
  w2eng_100_trim = Winsorize(w2eng_100, minval = 0, maxval = 100)
)
```

```{r Outcome variables}
d_cl <- d_cl |> mutate(
  w1ttl_100 = w1chn_100_trim + w1mat_100_trim + w1eng_100_trim,
  w2ttl_100 = w2chn_100_trim + w2mat_100_trim + w2eng_100_trim,
  w1avg_100 = w1ttl_100/3,
  w2avg_100 = w2ttl_100/3,
  w1cog_100 = stcog*100/20,
  w2cog_100 = w2cogscore*100/35,
  avg_dif = w2avg_100 - w1avg_100,
  cog_dif = w2cog_100 - w1cog_100,
  ) |>
  drop_na(avg_dif, cog_dif)
```

```{r Control variables}
d_cl <- d_cl |> mutate(
         child.age = 
           if_else(
             fall == 1, 
             2014 - a02a - a02b/12, 
             2014 - a02a - a02b/12 + 0.5),
         child.is.girl = if_else(stsex == 1, 0, 1), 
         child.has.rural.hukou = sthktype, 
         child.in.boarding.school = stbrd, 
         child.is.ethnic.minority = if_else(a03 != 1, 1, 0), 
         child.went.to.preschool = if_else(c01 == 1, 1, 0), 
         child.number.of.sibling = if_else(stonly == 1, 0, stsib),
         home.economic.resource = steco_5c, 
         home.has.internet = if_else(b13 == 2, 1, 0), 
         home.extracurricular.book = b12 
)  
```

```{r Parent education}
d_cl <- d_cl |> mutate(
  parent.education.level = 
    case_when(
      stprhedu == 1|stprhedu == 2 ~ 0,
      stprhedu == 3 ~ 1,
      stprhedu == 4|stprhedu == 5|stprhedu == 6 ~ 2,
      stprhedu == 7 ~ 3,
      stprhedu == 8 ~ 4,
      stprhedu == 9 ~ 5)
)
```

```{r Parent expectation}
d_cl <- d_cl |> mutate(
  parexp = 
    if_else(
      is.na(ba18),
        b31,
        ba18)
) |> mutate(
  parent.expectation = 
    case_when(
      parexp == 10 ~ 0,
      parexp == 1|parexp == 2 ~ 1,
      parexp == 3|parexp == 4|parexp == 5 ~ 2,
      parexp == 6 ~ 3,
      parexp == 7 ~ 4,
      parexp == 8|parexp == 9 ~ 5
    )
)
```

```{r Parent job}
d_cl <- d_cl |> mutate(
  b08ch = if_else(
    b08a==1|b08a==2|b08a==3|b08b==1|b08b==2|b08b==3,
    1,
    0
  ),
  b08pr = 
    if_else(be08==1|be08==2|be08==3, 1, 0)
)
d_cl <- d_cl |> mutate(
  parent.has.white.collar.job = 
    if_else(
      is.na(b08ch),
      b08pr,
      b08ch
    )
)
```

```{r Child aspiration}
d_cl <- d_cl |> mutate(
  child.education.aspiration = 
    case_when(
      c22 == 10 ~ 0,
      c22 == 1|c22 == 2 ~ 1,
      c22 == 3|c22 == 4|c22 == 5 ~ 2,
      c22 == 6 ~ 3,
      c22 == 7 ~ 4,
      c22 == 8|c22 == 9 ~ 5
    )
)
```

```{r Child health}
d_cl <- d_cl |> remove_labels() |>
  mutate(
  child.health = 
    if_else(
      is.na(a17),
        bd13,
        a17
    )
  )
```

```{ Child short sighted}
d_cl <- d_cl |> mutate(
  a15 = if_else(a15 == 3, 0, 1),
  bd14 = if_else(bd14 == 2, 0, 1)
  )
d_cl <- d_cl |> 
  mutate(child.short.sighted = 
           if_else(
             is.na(a15),
              bd14,
              a15
           ))

```

```{r Drop NA}
d_cl <- d_cl |>  drop_na(
  child.age, child.is.girl, child.has.rural.hukou, child.in.boarding.school, child.is.ethnic.minority, child.went.to.preschool, child.number.of.sibling, 
  #child.had.skipped.grade,
  #child.had.repeated.grade,
home.has.internet, home.extracurricular.book, home.economic.resource,
  #mother.education.level, 
  #father.education.level,
  parent.education.level,
  parent.expectation, 
  child.health,
  #child.short.sighted,
  child.education.aspiration)
```

## Treatment groups
At least one parent migrated (Control group as no parent migrated)
```{r Any parent migrated}
d_cl <- d_cl |> mutate(
  migr = if_else(w2ba0201==0 | w2ba0202==0, 1, 0) #Any parent became absent from household between baseline and endline survey
  ) |> drop_na(migr)
```

```{r Only mother migrated}
d_cl <- d_cl |> 
  mutate(
    migr_mo_only = 
      case_when(
        w2ba0201==1 & w2ba0202==0 ~ 1,
        migr == 0 ~ 0,
        .default = NA
      )
  )
```

```{r Only father migrated}
d_cl <- d_cl |> 
  mutate(
    migr_fa_only = 
      case_when(
        w2ba0201==0 & w2ba0202==1 ~ 1,
        migr == 0 ~ 0,
        .default = NA
      )
  )
```

```{r Both parents migrated}
d_cl <- d_cl |> 
  mutate(
    migr_both = 
      case_when(
        w2ba0201==0 & w2ba0202==0 ~ 1,
        migr == 0 ~ 0,
        .default = NA
      )
  )
```

## Descriptive statistics
```{r Cleaned data}
df <- d_cl |>
  select(
  migr, migr_mo_only, migr_fa_only, migr_both, schids, ctyids,
  w1cog_100, w2cog_100, w1avg_100, w2avg_100, cog_dif, avg_dif,
  child.age, child.number.of.sibling, child.is.girl, child.has.rural.hukou, child.in.boarding.school, #child.had.skipped.grade, child.had.repeated.grade, 
  child.is.ethnic.minority, child.went.to.preschool, parent.has.white.collar.job, parent.expectation, parent.education.level, home.has.internet,
  child.health, child.education.aspiration, 
  #mother.education.level, father.education.level, 
  home.extracurricular.book, home.economic.resource) 
```

```{r}
d_migr <- df |>
  mutate(`Parental Migration` = if_else((migr==1), 'Yes', 'No')) |>
  select(!migr)
```

```{r Data for treatment variable summary}
d_idpvar <- df |> 
  select(migr:migr_both) |>
  mutate(`Parental Migration` = if_else(migr==1, 'Yes', 'No')) |>
  mutate(
    `Parental Migration Type` = 
      case_when(
        migr==0 ~ 'Both Parents at Home',
        migr_mo_only==1 ~ 'Only Mother Migrated',
        migr_fa_only==1 ~ 'Only Father Migrated',
        migr_both==1 ~ 'Both Parents Migrated'
      )
) 
d_idpvar <- d_idpvar |> select(!migr & !migr_mo_only & !migr_fa_only & !migr_both)
```

```{r Treatment variable summary}
tab_idpvar <- st(
  data = d_idpvar,
  out = 'return'
)
tab_idpvar[tab_idpvar == 4204] <- ''
tab_idpvar
```

```{r Data for outcome variable summary}
d_dpvar <- d_migr |>
  select(`Parental Migration`, cog_dif, w1cog_100, w2cog_100, avg_dif, w1avg_100, w2avg_100) |>
  set_variable_labels(
    w1cog_100 ='Wave 1 score',
    w2cog_100 = 'Wave 2 score',
    w1avg_100 ='Wave 1 score',
    w2avg_100 = 'Wave 2 score',
    cog_dif = 'Cognitive test score difference',
    avg_dif = 'Academic exam score difference'
  )
```

```{r Outcome variable summary}
tab_dpvar <- 
  st(
  data = d_dpvar,
  summ = c('mean(x)', 'sd(x)'),
  digits = 2, 
  group = 'Parental Migration',
  group.test = list(
    star.cutoffs = c(0.001, 0.01, 0.05),
    format = 'p={pval}{stars}'),
  labels = T,
  out = 'return') 
tab_dpvar <- tab_dpvar[-c(1),]
tab_dpvar[tab_dpvar =='p=&lt0.001***']<-'p<0.001***'
tab_dpvar 
```

```{r Data for binary control variable summary}
d_ctvar_dit <- d_migr |>
  select(`Parental Migration`,child.is.girl, child.has.rural.hukou, child.in.boarding.school,
         #child.had.skipped.grade, child.had.repeated.grade, 
         child.is.ethnic.minority, child.went.to.preschool, parent.has.white.collar.job, home.has.internet) |> 
  mutate_all(as_factor)
```

```{r Binary control variable summary}
tab_ctvar_dit <- 
  st(
  data = d_ctvar_dit,
  digits = 2,
  group = 'Parental Migration',
  group.test = list(star.cutoffs = c(0.001, 0.01, 0.05)),
  out = "return"
  )
tab_ctvar_dit <- tab_ctvar_dit[-c(1),]
tab_ctvar_dit[tab_ctvar_dit == 3529|tab_ctvar_dit == 675] <- ''
tab_ctvar_dit 
```

```{r Data for categorical control variable summary}
d_ctvar_cnt <- d_migr |>
  select(`Parental Migration`, child.age, child.number.of.sibling, child.health, child.education.aspiration, 
  parent.education.level, parent.expectation, 
  home.extracurricular.book, home.economic.resource)
```

```{r Categorical control variable summary}
tab_ctvar_cnt <- d_ctvar_cnt |> st(
  digits = 2, 
  summ = c('mean(x)', 'sd(x)'),
  group = 'Parental Migration',
  group.test = list(star.cutoffs = c(0.001, 0.01, 0.05)),
  out = "return")
tab_ctvar_cnt <- tab_ctvar_cnt[-c(1),]
tab_ctvar_cnt
```

```{r Data for categorical control variable measurement}
d_ctvar_label <- d_ctvar_cnt|>
  select(!(`Parental Migration`:child.number.of.sibling)) |> 
  add_value_labels(
    parent.education.level = c(
     '0 = Primary school and below' = 0,
     '1 = Junior high school' = 1 ,
     '2 = Senior high school' = 2,
     '3 = Associate college' = 3,
     '4 = University' = 4,
     '5 = Postgraduate' = 5),
    child.education.aspiration = c(
     '0 = Does not matter' = 0,
     '1 = Junior high school' = 1,
     '2 = Senior high school' = 2,
     '3 = Associate college' = 3,
     '4 = University' = 4,
     '5 = Postgraduate' = 5),
    parent.expectation = c(
     '0 = Does not matter' = 0,
     '1 = Junior high school' = 1 ,
     '2 = Senior high school' = 2,
     '3 = Associate college' = 3,
     '4 = University' = 4,
     '5 = Postgraduate' = 5),
    home.economic.resource = c(
      '1 = Very poor' = 1,
      '2 = Somewhat poor' = 2,
      '3 = Moderate' = 3,
      '4 = Somewhat rich' = 4,
      '5 = Very rich' = 5),
    child.health = c(
      '1 = Very poor' = 1,
      '2 = Somewhat poor' = 2,
      '3 = Moderate' = 3,
      '4 = Good' = 4,
      '5 = Very good' = 5),
    home.extracurricular.book = c(
      '1 = Very few' = 1,
      '2 = Not many' = 2,
      '3 = Some' = 3,
      '4 = Quite a few' = 4,
      '5 = A great number' = 5),
    ) |>
  as_factor()
```

```{r Measurement for categorical control variable}
tab_var <- st(
  data = d_ctvar_label,
  out = 'return'
)
tab_var[tab_var == 4204] <- ''
tab_var
```

```{r Data for violin plots}
d_smr <- df |> 
  mutate(
    `Parental Migration` = 
      if_else(
        (migr==1), 'Yes', 'No')
    ) |>
  mutate(
    `Parental Migration Type` = 
      case_when(
        migr==0 ~ 'Both Parents at Home',
        migr_mo_only==1 ~ 'Only Mother Migrated',
        migr_fa_only==1 ~ 'Only Father Migrated',
        migr_both==1 ~ 'Both Parents Migrated'
      )
  )
d_smr <- d_smr |> select(!migr_mo_only & !migr_fa_only & !migr_both)
```

```{r Violin plot 1}
set.seed(1)
fig_cog <- d_smr |>
  ggplot(aes(x=`Parental Migration`, y=cog_dif, fill=`Parental Migration`)) + 
    geom_violin()+
    xlab('Parental Migration') +
    theme(legend.position="none") +
    ylab('Cognitive Test Score Difference') +
    stat_summary(
      fun = "mean",
      geom = "crossbar", 
      size = 0.4,
      width = 0.5) +
  scale_fill_manual(values = alpha(c("green", "red"), .4))
fig_cog
```

```{r Violin plot 2}
set.seed(1)
fig_ttl <-  d_smr |>
  ggplot(aes(x=`Parental Migration`, y=avg_dif, fill=`Parental Migration`)) + 
    geom_violin()+
    xlab('Parental Migration') +
    theme(legend.position="none") +
    ylab('Academic Exam Score Difference') +
    stat_summary(
      fun = "mean",
      geom = "crossbar", 
      size = 0.4,
      width = 0.5) +
  scale_fill_manual(values = alpha(c("green", "red"), .4))
fig_ttl
```

```{r Violin plot 3}
set.seed(1)
fig_cog_by_migration_type <- d_smr |>
  ggplot(aes(x=`Parental Migration Type`, y=cog_dif, fill = `Parental Migration Type`)) + 
    geom_violin()+
    xlab('Parental Migration Type') +
    theme(legend.position="none") +
    ylab('Cognitive Test Score Difference') +
    stat_summary(
      fun = "mean",
      geom = "crossbar", 
      width = 0.5)+
  scale_fill_manual(values = alpha(c("green", "tomato1", "tomato2", "tomato3"), .4))
fig_cog_by_migration_type
```

```{r Violin plot 4}
set.seed(1)
fig_ttl_by_migration_type <- d_smr |>
  ggplot(aes(x=`Parental Migration Type`, y=avg_dif, fill = `Parental Migration Type`)) + 
    geom_violin()+
    xlab('Parental Migration Type') +
    theme(legend.position="none") +
    ylab('Academic Exam Score Difference') +
    stat_summary(
      fun = "mean",
      geom = "crossbar", 
      width = 0.5) +
  scale_fill_manual(values = alpha(c("green", "tomato1", "tomato2", "tomato3"), .4))
fig_ttl_by_migration_type
```

# Econometric models
## DID without covariates
```{r Unadjusted regression on cognitive score}
did_cog <- feols(cog_dif ~ migr + w1cog_100 | schids, data = df) 
did_cog_m_only <- feols(cog_dif ~ migr_mo_only + w1cog_100 | schids, data = df) 
did_cog_f_only <- feols(cog_dif ~ migr_fa_only + w1cog_100 | schids, data = df) 
did_cog_both <- feols(cog_dif ~ migr_both + w1cog_100 | schids, data = df)
```

```{r Unadjusted regression on exam grade}
did_ttl <- feols(avg_dif ~ migr + w1avg_100 | schids, data = df) 
did_ttl_m_only <- feols(avg_dif ~ migr_mo_only + w1avg_100 | schids, data = df) 
did_ttl_f_only <- feols(avg_dif ~ migr_fa_only + w1avg_100 | schids, data = df) 
did_ttl_both <- feols(avg_dif ~ migr_both + w1avg_100 | schids, data = df)
```

```{r}
models_cog <- list(
'Any Parent' = did_cog,
'Only Mother' = did_cog_m_only,
'Only Father' = did_cog_f_only,
'Both Parents' = did_cog_both) 
```

```{r}
models_ttl <- list(
'Any Parent' = did_ttl,
'Only Mother' = did_ttl_m_only,
'Only Father' = did_ttl_f_only,
'Both Parents' = did_ttl_both) 
```

```{r Info for regression tables}
stars <- c('*' = .05, '**' = .01, '***' = .001)
coef_rename <- c("migr" = "Migration", "migr_mo_only" = "Migration", "migr_mo" = "Migration", "migr_fa" = "Migration", "migr_fa_only" = "Migration", "migr_both" = "Migration", "w1cog_100" = "Baseline.score", "w1avg_100" = "Baseline.score")
title_cog <- 'Parental migration’s effect on children’s cognitive test score'
title_ttl <- 'Parental migration’s effect on children’s academic exam score'
gof_omit <- 'R2 Within|R2 Pseudo|RMSE|AIC|BIC|Log.Lik.|Std.Errors'
note_unadj <- 'No control variables included in these models'
```

```{r}
models_cog <- list(
'Any Parent' = did_cog,
'Only Mother' = did_cog_m_only,
'Only Father' = did_cog_f_only,
'Both Parents' = did_cog_both) 
```

```{r}
models_ttl <- list(
'Any Parent' = did_ttl,
'Only Mother' = did_ttl_m_only,
'Only Father' = did_ttl_f_only,
'Both Parents' = did_ttl_both) 
```

```{r Table for unadjusted regression on cognitive score}
table_cog <- models_cog |> msummary(
  stars =  stars,
  fmt = 2,
  title = title_cog,
  gof_omit = gof_omit,
  coef_rename =  coef_rename,
  notes = note_unadj,
  output = 'latex')
```

```{r Table for unadjusted regression on exam grade}
table_ttl <- models_ttl |> msummary(
  stars =  stars,
  fmt = 2,
  title = title_ttl,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  notes = note_unadj,
  output = 'latex')
```

## DID with covariates
$$\Delta score_{i,s} = \alpha + \beta \cdot migr_{i,s} + \gamma \cdot FE_{s} + \lambda \cdot score_{i,s;base} + \theta \cdot X_{i,s} + \varepsilon_{i,s}$$

```{r Adjusted regression on cognitive score}
did_adj_cog <- feols(cog_dif ~ migr + w1cog_100  + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + parent.education.level + home.economic.resource + home.extracurricular.book  + home.has.internet | schids, data = df) 
did_adj_cog_m_only <- feols(cog_dif ~ migr_mo_only + w1cog_100  + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + parent.education.level + home.economic.resource + home.extracurricular.book  + home.has.internet | schids, data = df) 
did_adj_cog_f_only <- feols(cog_dif ~ migr_fa_only + w1cog_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + parent.education.level + home.economic.resource + home.extracurricular.book  + home.has.internet | schids, data = df) 
did_adj_cog_both <- feols(cog_dif ~ migr_both + w1cog_100  + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + parent.education.level + home.economic.resource + home.extracurricular.book  + home.has.internet | schids, data = df)
```

```{r Adjusted regression on exam grade}
did_adj_ttl <- feols(avg_dif ~ migr + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + parent.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = df) 
did_adj_ttl_m_only <- feols(avg_dif ~ migr_mo_only + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + parent.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = df) 
did_adj_ttl_f_only <- feols(avg_dif ~ migr_fa_only + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + parent.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = df) 
did_adj_ttl_both <- feols(avg_dif ~ migr_both + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + parent.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = df)
```

```{r}
models_adj_cog <- list(
'Any Parent' = did_adj_cog,
'Only Mother' = did_adj_cog_m_only,
'Only Father' = did_adj_cog_f_only,
'Both Parents' = did_adj_cog_both) 
```

```{r}
models_adj_ttl <- list(
'Any Parent' = did_adj_ttl,
'Only Mother' = did_adj_ttl_m_only,
'Only Father' = did_adj_ttl_f_only,
'Both Parents' = did_adj_ttl_both)
```

```{r Table for adjusted regression on cognitive score}
table_adj_cog <- models_adj_cog |> msummary(
  stars =  stars,
  fmt = 2,
  title = title_cog,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  output = 'latex'
  )
```

```{r Table for adjusted regression on exam grade}
table_adj_ttl <- models_adj_ttl |> msummary(
  stars =  stars,
  fmt = 2,
  title = title_ttl,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  output = 'latex'
  )
```

```{}
table_cog
table_ttl
table_adj_cog
table_adj_ttl
```

## Matching
```{r Match on any parent migration}
m.out <- matchit(
  migr ~ w1cog_100 + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.in.boarding.school + child.is.ethnic.minority + child.education.aspiration + parent.expectation + parent.has.white.collar.job + parent.education.level + home.economic.resource + home.extracurricular.book + home.has.internet, 
  data = df, 
  method = "full", 
  distance = "glm"
  )
```

```{r}
dm <- match.data(m.out)
```

```{r Match on mother-only migration}
d_mo_only <- d_cl |> drop_na(migr_mo_only)
m.out_mo_only <- 
  matchit(migr_mo_only ~ w1cog_100 + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.in.boarding.school + child.is.ethnic.minority + child.education.aspiration + parent.expectation + parent.has.white.collar.job + parent.education.level + home.economic.resource + home.extracurricular.book + home.has.internet, 
  data = d_mo_only, 
  method = "full",
  distance = "glm"
  )
```

```{r}
dm_mo_only <- match.data(m.out_mo_only)
```

```{r Match on father-only migration}
d_fa_only <- d_cl |> drop_na(migr_fa_only)
m.out_fa_only <- matchit(
  migr_fa_only ~ w1cog_100 + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.in.boarding.school + child.is.ethnic.minority + child.education.aspiration + parent.expectation + parent.has.white.collar.job + parent.education.level + home.economic.resource + home.extracurricular.book + home.has.internet,
  data = d_fa_only,
  method = "full",
  distance = "glm"
  )
```

```{r}
dm_fa_only <- match.data(m.out_fa_only)
```

```{r Match on both parents migration}
d_both <- d_cl |> drop_na(migr_both) 
m.out_both <- matchit(
  migr_both ~ w1cog_100 + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.in.boarding.school + child.is.ethnic.minority + child.education.aspiration + parent.expectation + parent.has.white.collar.job + parent.education.level +  home.economic.resource + home.extracurricular.book + home.has.internet, 
  data = d_both, 
  method = "full",
  distance = "glm"
  )
```

```{r}
dm_both <- match.data(m.out_both)
```

```{r}
summary(m.out)
summary(m.out_mo_only)
summary(m.out_fa_only)
summary(m.out_both)
```

## Balance Assessment
```{r ASMD plot any-parent}
m.out |> love.plot(
  threshold = c(m = .1),
  binary = "std", 
  limits = c(0, 0.55),
  abs = T
  )
```

```{r ASMD plot only-mother}
m.out_mo_only |> love.plot(
  threshold = c(m = .1),
  binary = "std", 
  limits = c(0, 0.55),
  abs = T
  )
```

```{r ASMD plot only-father}
m.out_fa_only |> love.plot(
  threshold = c(m = .1),
  binary = "std", 
  limits = c(0, 0.55),
  abs = T
  )
```

```{r ASMD plot both-parents}
m.out_both |> love.plot(
  threshold = c(m = .1),
  binary = "std", 
  limits = c(0, 0.55),
  abs = T
  )
```

## Regression with matching

```{r Unadjusted regression on cognitive score with matching}
didm_cog <- feols(cog_dif ~ migr + w1cog_100 | schids, data = dm, weights = dm$weights)
didm_cog_mo_only <- feols(cog_dif ~ migr_mo_only + w1cog_100 | schids, data = dm_mo_only, weights = dm_mo_only$weights)
didm_cog_fa_only <- feols(cog_dif ~ migr_fa_only + w1cog_100 | schids, data = dm_fa_only, weights = dm_fa_only$weights)
didm_cog_both <- feols(cog_dif ~ migr_both + w1cog_100 | schids, data = dm_both, weights = dm_both$weights)
```

```{r Unadjusted regression on exam score with matching}
didm_ttl <- feols(avg_dif ~ migr + w1avg_100 | schids, data = dm, weights = dm$weights)
didm_ttl_mo_only <- feols(avg_dif ~ migr_mo_only + w1avg_100 | schids, data = dm_mo_only, weights = dm_mo_only$weights)
didm_ttl_fa_only <- feols(avg_dif ~ migr_fa_only + w1avg_100 | schids, data = dm_fa_only, weights = dm_fa_only$weights)
didm_ttl_both <- feols(avg_dif ~ migr_both + w1avg_100 | schids, data = dm_both, weights = dm_both$weights)
```

```{r Adjusted regression on cognitive score with matching}
didm_adj_cog <- feols(cog_dif ~ migr + w1cog_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + parent.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = dm, weights = dm$weights)
didm_adj_cog_mo_only <- feols(cog_dif ~ migr_mo_only + w1cog_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + parent.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = dm_mo_only, weights = dm_mo_only$weights)
didm_adj_cog_fa_only <- feols(cog_dif ~ migr_fa_only + w1cog_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + parent.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = dm_fa_only, weights = dm_fa_only$weights)
didm_adj_cog_both <- feols(cog_dif ~ migr_both + w1cog_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + parent.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = dm_both, weights = dm_both$weights)
```

```{r Adjusted regression on exam score with matching}
didm_adj_ttl <- feols(avg_dif ~ migr + w1cog_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + parent.education.level + home.economic.resource + home.extracurricular.book + home.has.internet | schids, data = dm, weights = dm$weights)
didm_adj_ttl_mo_only <- feols(avg_dif ~ migr_mo_only + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + parent.education.level + home.economic.resource + home.extracurricular.book + home.has.internet  | schids, data = dm_mo_only, weights = dm_mo_only$weights)
didm_adj_ttl_fa_only <- feols(avg_dif ~ migr_fa_only + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + parent.education.level + home.economic.resource + home.extracurricular.book + home.has.internet  | schids, data = dm_fa_only, weights = dm_fa_only$weights)
didm_adj_ttl_both <- feols(avg_dif ~ migr_both + w1avg_100 + child.age + child.is.girl + child.health + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.in.boarding.school + child.education.aspiration + child.is.ethnic.minority + parent.expectation + parent.has.white.collar.job + parent.education.level + home.economic.resource + home.extracurricular.book + home.has.internet  | schids, data = dm_both, weights = dm_both$weights)
```

```{r Info for regression table with matching}
title_cog_matching <- 'Parental migration’s effect on children’s cognitive test score, estimated with matching'
title_ttl_matching <- 'Parental migration’s effect on children’s academic exam score, estimated with matching'
note_matching <- "Optimal full matching by propensity score"
```

```{r}
models_m_cog <- list(
'Any Parent' = didm_cog,
'Only Mother' = didm_cog_mo_only,
'Only Father' = didm_cog_fa_only,
'Both Parents' = didm_cog_both) 
```

```{r}
models_m_ttl <- list(
'Any Parent' = didm_ttl,
'Only Mother' = didm_ttl_mo_only,
'Only Father' = didm_ttl_fa_only,
'Both Parents' = didm_ttl_both) 
```

```{r}
models_m_adj_cog <- list(
'Any Parent' = didm_adj_cog,
'Only Mother' = didm_adj_cog_mo_only,
'Only Father' = didm_adj_cog_fa_only,
'Both Parents' = didm_adj_cog_both) 
```

```{r}
models_m_adj_ttl <- list(
'Any Parent' = didm_adj_ttl,
'Only Mother' = didm_adj_ttl_mo_only,
'Only Father' = didm_adj_ttl_fa_only,
'Both Parents' = didm_adj_ttl_both) 
```

```{r Table for unadjusted regression on cognitive score with matching}
table_m_cog <- models_m_cog |> msummary(
  stars =  stars,
  fmt = 2,
  title = title_cog_matching,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  notes = note_matching,
  output = 'latex'
  )
```

```{r Table for unadjusted regression on exam grade with matching}
table_m_ttl <- models_m_ttl |> msummary(
  stars =  stars,
  fmt = 2,
  title = title_ttl_matching,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  notes = note_matching,
  output = 'latex'
)
```

```{r Table for adjusted regression on cognitive score with matching}
table_m_adj_cog <- models_m_adj_cog |> msummary(
  stars =  stars,
  fmt = 2,
  title = title_cog_matching,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  notes = note_matching,
  output = 'latex'
  )
```

```{r Table for adjusted regression on exam grade with matching}
table_m_adj_ttl <- models_m_adj_ttl |> msummary(
  stars =  stars,
  fmt = 2,
  title = title_ttl_matching,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  notes = note_matching,
  output = 'latex'
 )
```

```{}
table_m_cog
table_m_ttl
table_m_adj_cog
table_m_adj_ttl
```

```{ save image}
save.image('.RData')
```